---
title: "RT-qPCR of pPIR1-mCh-tPIR1mod (2 reps), pPGK1-mCh-tPIR1mod (2 reps) and pSRO9-mCh-tPIR1mod (2 reps)"
author: "Jamie Auxillos"
date: "06/05/2020"
output:
  html_document:
    toc: yes
    #toc_depth: 2
---

```{r setup,warning=FALSE,message=FALSE,echo=FALSE}

## knitr options for report generation
knitr::opts_chunk$set(warning=FALSE,message=FALSE,echo=TRUE,cache=FALSE)


library(dplyr)
library(tibble)
library(stringr)
library(readr)
library(cowplot)
library(tidyqpcr)
library(here)

 # set default theme for graphics
source(here("raw_data_analysis/code/shared_figure_formatting.R"))
```

```{r label_plates, echo=FALSE}

# list Targets (Primer sets)
target_id <- c("mCh-7", "URA3-ORF", "PIR1-ORF", "PGK1-ORF")
target_id_values <-factor(rep(target_id,levels=target_id))


# labelling of SampleID attributes
strain_levels_PIR1 <- c("pPIR1-mCherry-tPIR1_WT","pPIR1-mCherry-tPIR1_modA","pPIR1-mCherry-tPIR1_modB","pPIR1-mCherry-tPIR1_modC", "pPIR1-mCherry-tPIR1_modD", "pPIR1-mCherry-tPIR1_modE", "pPIR1-mCherry-tPIR1_modF", "pPIR1-mCherry-tPIR1_modG")
strain_values_PIR1 <- factor(rep(strain_levels_PIR1,each=3),levels=strain_levels_PIR1)
strain_levels_PGK1 <- c("pPGK1-mCherry-tPIR1_WT","pPGK1-mCherry-tPIR1_modA","pPGK1-mCherry-tPIR1_modB","pPGK1-mCherry-tPIR1_modC", "pPGK1-mCherry-tPIR1_modD", "pPGK1-mCherry-tPIR1_modE", "pPGK1-mCherry-tPIR1_modF", "pPGK1-mCherry-tPIR1_modG")
strain_values_PGK1 <- factor(rep(strain_levels_PGK1,each=3),levels=strain_levels_PGK1)
strain_levels_SRO9 <- c("pSRO9-mCherry-tPIR1_WT","pSRO9-mCherry-tPIR1_modA","pSRO9-mCherry-tPIR1_modB","pSRO9-mCherry-tPIR1_modC", "pSRO9-mCherry-tPIR1_modD", "pSRO9-mCherry-tPIR1_modE","pSRO9-mCherry-tPIR1_modF", "pSRO9-mCherry-tPIR1_modG")
strain_values_SRO9 <- factor(rep(strain_levels_SRO9,each=3),levels=strain_levels_SRO9)
bio_rep_levels_A <- c("bio_rep1","bio_rep2","bio_rep3")
bio_rep_values_A <- factor(rep(bio_rep_levels_A,times=8),levels=bio_rep_levels_A)
bio_rep_levels_B <- c("bio_rep4","bio_rep5","bio_rep6")
bio_rep_values_B <- factor(rep(bio_rep_levels_B,times=8),levels=bio_rep_levels_B)

# Attribute labels to each row and column for each plate (pPIR1 and pPGK1 plates)
colkey_PIR1_1 <- tibble(well_col=1:24, strain=strain_values_PIR1, bio_rep=bio_rep_values_A) 
colkey_PIR1_2 <- tibble(well_col=1:24, strain=strain_values_PIR1, bio_rep=bio_rep_values_B) 
colkey_PGK1_1 <- tibble(well_col=1:24, strain=strain_values_PGK1, bio_rep=bio_rep_values_A) 
colkey_PGK1_2 <- tibble(well_col=1:24, strain=strain_values_PGK1, bio_rep=bio_rep_values_B) 
colkey_SRO9_1 <- tibble(well_col=1:24, strain=strain_values_SRO9, bio_rep=bio_rep_values_A) 
colkey_SRO9_2 <- tibble(well_col=1:24, strain=strain_values_SRO9, bio_rep=bio_rep_values_B) 
rowkey <- create_rowkey_4_in_16(target_id=target_id_values) 


# Combine all the information into 4x plate plans
plateplan_PIR1_1 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_PIR1_1%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modG","modF","modE","modD","modC","modB","modA","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_PIR1_2 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_PIR1_2%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modG","modF","modE","modD","modC","modB","modA","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_PGK1_1 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_PGK1_1%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modG","modF","modE","modD","modC","modB","modA","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_PGK1_2 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_PGK1_2%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modG","modF","modE","modD","modC","modB","modA","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_SRO9_1 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_SRO9_1%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modG","modF","modE","modD","modC","modB","modA","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_SRO9_2 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_SRO9_2%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modG","modF","modE","modD","modC","modB","modA","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))
```

## Load data and attribute sample information
```{r load_plates,dependson="label_plates",results="show"}
# read my plates


plates_PIR1_1 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPIR1-tPIR1/JA_20210101-pPIR1-tPIR1_mod-n1_repeatloading-ct.txt")) %>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_PIR1_2)

plates_PIR1_2 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPIR1-tPIR1/JA_20210101-pPIR1-tPIR1mod-n2-ct.txt")) %>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_PIR1_1)

plates_PGK1_1 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tPIR1/JA_20201212-pPGK1-tPIR1mod-n1-ct.txt")) %>%
    mutate(exp_rep = 1, exp_rep=factor(exp_rep)) %>% 
    left_join(plateplan_PGK1_1)

plates_PGK1_2 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tPIR1/JA_20201223-pPGK1-tPIR1mod-n2-ct.txt")) %>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_PGK1_2)

plates_SRO9_1 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tPIR1/JA_20210101-pSRO9-tPIR1mod-n1-ct.txt")) %>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_SRO9_1)

plates_SRO9_2 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tPIR1/JA_20210101-pSRO9-tPIR1mod-n2-ct.txt")) %>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_SRO9_2)


plates <- bind_rows(plates_PIR1_1,plates_PIR1_2,plates_PGK1_1,plates_PGK1_2, plates_SRO9_1, plates_SRO9_2) 

summary(plates)
```


### Plotting unnormalised data for pPIR1 strains
Here we are plotting the unnormalised qPCR data for the constructs containing pPIR1 driving the expression of mCherry. All -RT samples appear to have high cq values and all +RT samples have cq values below 20 cycles. There is however a clear difference in the cq values between experimental replicates (Experimental replicate 1 -Biorep 1, 2, 3)(Experimental replicate 2 -Biorep 4, 5, 6). Experimental replicate 2 seems to have lower cq values across all primers and strains than experimental replicate 1.  

```{r unnorm_pPIR1,fig.height=5,fig.width=10, echo=FALSE}
plates_PIR1 <- plates%>% filter(strain %in% c("pPIR1-mCherry-tPIR1_WT","pPIR1-mCherry-tPIR1_modA","pPIR1-mCherry-tPIR1_modB",
                                           "pPIR1-mCherry-tPIR1_modC","pPIR1-mCherry-tPIR1_modD","pPIR1-mCherry-tPIR1_modE","pPIR1-mCherry-tPIR1_modF","pPIR1-mCherry-tPIR1_modG"))%>%
  mutate(strain = factor(strain,levels=c("pPIR1-mCherry-tPIR1_WT","pPIR1-mCherry-tPIR1_modA","pPIR1-mCherry-tPIR1_modB",
                                           "pPIR1-mCherry-tPIR1_modC","pPIR1-mCherry-tPIR1_modD","pPIR1-mCherry-tPIR1_modE","pPIR1-mCherry-tPIR1_modF","pPIR1-mCherry-tPIR1_modG")))

ggplot(data=plates_PIR1, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.6, alpha=0.7) +
    labs(y="Quantification cycle (cq)",x="Primers") +
    scale_shape_manual(values=c(16,23)) +
    scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
    facet_wrap(~strain,ncol=4) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
         # legend.position=c(0.85,0.01))
```


### Plotting unnormalised data for pPGK1 strains
Here we are plotting the unnormalised qPCR data for the constructs containing pPIR1 driving the expression of mCherry. All -RT samples appear to have high cq values and all +RT samples have cq values below 20 cycles. There is little difference between the experimental replicates.

```{r unnorm_pPGK1,fig.height=5,fig.width=10, echo=FALSE}

plates_PGK1 <- plates%>% filter(strain %in% c("pPGK1-mCherry-tPIR1_WT","pPGK1-mCherry-tPIR1_modA","pPGK1-mCherry-tPIR1_modB",
                                           "pPGK1-mCherry-tPIR1_modC","pPGK1-mCherry-tPIR1_modD","pPGK1-mCherry-tPIR1_modE","pPGK1-mCherry-tPIR1_modF","pPGK1-mCherry-tPIR1_modG"))%>%
  mutate(strain = factor(strain,levels=c("pPGK1-mCherry-tPIR1_WT","pPGK1-mCherry-tPIR1_modA","pPGK1-mCherry-tPIR1_modB",
                                           "pPGK1-mCherry-tPIR1_modC","pPGK1-mCherry-tPIR1_modD","pPGK1-mCherry-tPIR1_modE","pPGK1-mCherry-tPIR1_modF","pPGK1-mCherry-tPIR1_modG")))
ggplot(data=plates_PGK1, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.6, alpha=0.7) +
    labs(y="Quantification cycle (cq)",x="Primers") +
    scale_shape_manual(values=c(16,23)) +
    scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
    facet_wrap(~strain,ncol=4) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
         # legend.position=c(0.85,0.01))
```

### Plotting unnormalised data for pSRO9 strains
Here we are plotting the unnormalised qPCR data for the constructs containing pPIR1 driving the expression of mCherry. All -RT samples appear to have high cq values and all +RT samples have cq values below 20 cycles. There is little difference between the experimental replicates.
```{r unnorm_pSRO9,fig.height=5,fig.width=10, echo=FALSE}

plates_SRO9 <- plates%>% filter(strain %in% c("pSRO9-mCherry-tPIR1_WT","pSRO9-mCherry-tPIR1_modA","pSRO9-mCherry-tPIR1_modB",
                                           "pSRO9-mCherry-tPIR1_modC","pSRO9-mCherry-tPIR1_modD","pSRO9-mCherry-tPIR1_modE","pSRO9-mCherry-tPIR1_modF","pSRO9-mCherry-tPIR1_modG"))%>%
  mutate(strain = factor(strain,levels=c("pSRO9-mCherry-tPIR1_WT","pSRO9-mCherry-tPIR1_modA","pSRO9-mCherry-tPIR1_modB",
                                           "pSRO9-mCherry-tPIR1_modC","pSRO9-mCherry-tPIR1_modD","pSRO9-mCherry-tPIR1_modE","pSRO9-mCherry-tPIR1_modF","pSRO9-mCherry-tPIR1_modG")))
ggplot(data=plates_SRO9, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.6, alpha=0.7) +
    labs(y="Quantification cycle (cq)",x="Primers") +
    scale_shape_manual(values=c(16,23)) +
    scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
    facet_wrap(~strain,ncol=4) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
          #legend.position=c(0.85,0.01))
```


### Plotting unnormalised data for control strains (pPIR1-mCherry-tPIR1_WT)


```{r unnorm_POT_pPIR1, fig.height=3.5,fig.width=5.5, echo=FALSE}

plates_POT_PIR1 <- bind_rows(plates_PIR1_1,plates_PIR1_2)%>% filter(strain %in% c("pPIR1-mCherry-tPIR1_WT"))

ggplot(data=plates_POT_PIR1, aes(x=target_id,y=cq)) +
  geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.5, alpha=0.7) +
  labs(title="Controls for pPIR1 strains - \n Comparing between experimental replicates",y="Quantification cycle (cq)",x="Primers") +
  ylim(0,40)+
  scale_shape_manual(values=c(16,23)) +
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  facet_wrap(~strain,ncol=2)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
```


### Plotting unnormalised data for control strains (pPGK1-mCherry-tPIR1_WT)


```{r unnorm_POT_pPGK1, fig.height=3.5,fig.width=5.5, echo=FALSE}

plates_POT_PGK1 <- bind_rows(plates_PGK1_1,plates_PGK1_2)%>% filter(strain %in% c("pPGK1-mCherry-tPIR1_WT"))

ggplot(data=plates_POT_PGK1, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.5, alpha=0.7) +
  labs(title="Controls for pPGK1 strains - \n Comparing between experimental replicates",y="Quantification cycle (cq)",x="Primers") +
  ylim(0,40)+
  scale_shape_manual(values=c(16,23)) +
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  facet_wrap(~strain,ncol=2)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
```

### Plotting unnormalised data for control strains (pSRO9-mCherry-tPIR1_WT)

```{r unnorm_POT_pSRO9, fig.height=3.5,fig.width=5.5, echo=FALSE}

plates_POT_SRO9 <- bind_rows(plates_SRO9_1,plates_SRO9_2)%>% filter(strain %in% c("pSRO9-mCherry-tPIR1_WT"))

ggplot(data=plates_POT_SRO9, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.5, alpha=0.7) +
  labs(title="Controls for pSRO9 strains - \n Comparing between experimental replicates",y="Quantification cycle (cq)",x="Primers") +
  ylim(0,40)+
  scale_shape_manual(values=c(16,23)) +
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  facet_wrap(~strain,ncol=2)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
```

## Loading amplification and melt curve pPIR1 data

```{r load_amp-melt_PIR1,dependson="label_plates", results="show"}

platecurve_PIR1_1 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPIR1-tPIR1/JA_20210101-pPIR1-tPIR1_mod-n1_repeatloading.txt"),skip=2,
                      col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_PIR1_1)%>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep))

platecurve_PIR1_2 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPIR1-tPIR1/JA_20210101-pPIR1-tPIR1mod-n2.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_PIR1_2)%>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep))

platecurve_PIR1 <- bind_rows(platecurve_PIR1_1, platecurve_PIR1_2)

platesamp_PIR1  <- platecurve_PIR1 %>% filter(program_no == 2)
platesmelt_PIR1 <- platecurve_PIR1 %>% filter(program_no != 2) %>% getdRdTall() %>% filter(temperature >= 61)  
```

### Amplification Curves of strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, PIR1-ORF, URA3-ORF) (pPIR1)

```{r plotamp_Rep_PIR1,fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesamp_PIR1,
       aes(x=cycle,y=fluor_signal,linetype=prep_type,group=well, color=exp_rep)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line() +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    expand_limits(y=0) + 
    labs(title="Amplification curves of strains (bio_rep 1-6) in pPIR1 experiments") + panel_border()

```



### Melt Curves ofstrains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, PIR1-ORF, URA3-ORF) (pPIR1)
  

```{r plotmelt_Rep_PIR1,dependson="load_amp",fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesmelt_PIR1 ,
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of POT1-ccdB strains (bio_rep 1-6) in pPIR1 experiments") + panel_border()
```


### Loading amplification and melt curve pPGK1 data

```{r load_amp-melt_P1,dependson="label_plates"}

platecurve_PGK1_1 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tPIR1/JA_20201212-pPGK1-tPIR1mod-n1.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_PGK1_1)%>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep))

platecurve_PGK1_2 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tPIR1/JA_20201223-pPGK1-tPIR1mod-n2.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_PGK1_2)%>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep))

platecurve_PGK1 <- bind_rows(platecurve_PGK1_1, platecurve_PGK1_2)

platesamp_PGK1  <- platecurve_PGK1 %>% filter(program_no == 2)
platesmelt_PGK1 <- platecurve_PGK1 %>% filter(program_no != 2) %>% getdRdTall() %>% filter(temperature >= 61)  
```

### Amplification Curves of strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, PIR1-ORF, URA3-ORF) (pPGK1)

```{r plotamp_Rep_PGK1,dependson="load_amp",fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesamp_PGK1,
       aes(x=cycle,y=fluor_signal,linetype=prep_type,group=well, color=exp_rep)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line() +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    expand_limits(y=0) + 
    labs(title="Amplification curves of strains (bio_rep 1-6) in pPGK1 experiments") + panel_border()
```

### Melt Curves of pPGK1-mCherry-tPIR1_WT strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, PIR1-ORF, URA3-ORF) (pPGK1)

```{r plotmelt_Rep_PGK1,fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesmelt_PGK1,
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of strains (bio_rep 1-6) in pPGK1 experiments") + panel_border()
```

### Loading amplification and melt curve pSRO9 data

```{r load_amp-melt_SRO9,dependson="label_plates"}

platecurve_SRO9_1 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tPIR1/JA_20210101-pSRO9-tPIR1mod-n1.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_SRO9_1)%>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep))

platecurve_SRO9_2 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tPIR1/JA_20210101-pSRO9-tPIR1mod-n2.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_SRO9_2)%>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep))

platecurve_SRO9 <- bind_rows(platecurve_SRO9_1, platecurve_SRO9_2)

platesamp_SRO9  <- platecurve_SRO9 %>% filter(program_no == 2)
platesmelt_SRO9 <- platecurve_SRO9 %>% filter(program_no != 2) %>% getdRdTall() %>% filter(temperature >= 61)  
```

### Amplification Curves of strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, PIR1-ORF, URA3-ORF) (pSRO9)


```{r plotamp_Rep_SRO9,dependson="load_amp",fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesamp_SRO9,
       aes(x=cycle,y=fluor_signal,linetype=prep_type,group=well, color=exp_rep)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line() +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    expand_limits(y=0) + 
    labs(title="Amplification curves of strains (bio_rep 1-6) in pSRO9 experiments") + panel_border()
```

### Melt Curves of POT1-ccdB and pPGK1-mCherry-tPIR1_WT strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, PIR1-ORF, URA3-ORF) (SRO9)

```{r plotmelt_Rep_SRO9,fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesmelt_SRO9,
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of POT1-ccdB strains (bio_rep 1-6) in pSRO9 experiments") + panel_border()
```


## 1) Delta cq calculation
norm_function changed to **mean**

```{r deltacq_norm, echo=TRUE, fig.height=6, fig.width=9}

# platesnorm first excludes pPGK1-mCherry-tPIR1_WT_bio_rep2, and all POT1-ccdB strains, then takes all +RT results, then normalises the cq values by the median of normtarget_ids, then filters cq values for mCh-7
platesnorm <- plates  %>% 
        filter(prep_type=="+RT") %>%    
        calculate_deltacq_bysampleid(ref_target_ids = c("PGK1-ORF", "RPS3-ORF", "URA3-ORF"), norm_function = mean)

platesnorm_summarise <- platesnorm%>%
                    group_by(strain, target_id, bio_rep, exp_rep)%>%
                    summarize(deltacq = median(delta_cq, na.rm=TRUE),
                              rel_abund_deltacq = (2^-deltacq), na.rm=FALSE)%>% 
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modG","modF","modE","modD","modC","modB","modA","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE)



ggplot(platesnorm, aes(UTR3,delta_cq))+
  geom_point(aes(color=bio_rep),position=position_dodge(width = 0.85),size=1.2, alpha=0.7)+
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  labs(y="delta cq", x="3'UTR-terminators")+
  facet_wrap(~target_id)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5))+
  geom_hline(yintercept = 0, color = "black", linetype= 'dotted', size=1) 
```

### 1) Exporting analysed delta cq values data
```{r deltacq_normdata, echo=TRUE}
#exporting platesnorm dataframe
write.csv(platesnorm_summarise,here("raw_data_analysis/data/norm_qpcr/motif_context_dependence/pPIR1_pPGK1_pSRO9_tPIR1_deltacq_platesnorm_summarise.csv"), row.names = FALSE)

```


```{r plot_deltacq, echo=TRUE, fig.height=5, fig.width=9}

ggplot(platesnorm %>% filter(target_id %in% 'mCh-7'), aes(UTR3,delta_cq))+
  geom_point(aes(color=exp_rep),position=position_dodge(width = 0.85),size=1.2, alpha=1)+
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  labs(y="delta cq", x="3'UTR-terminators")+
  facet_wrap(~promoter)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5))+
  geom_hline(yintercept = 0, color = "black", linetype= 'dotted', size=1) 
```

## 2) Delta-delta cq and RNA abundance calculation (one WT mean per exp_rep per promoter)
```{r delta-deltacq_norm_PIR1_exp_rep, echo=TRUE}

# Extracts the pPIR1 strains
platesnorm_pPIR1_exp_rep <- platesnorm %>% filter(promoter %in% "pPIR1", target_id %in% "mCh-7")
  
# Calculates the mean cq value of pPIR1-mCherry-tPIR1_WT Samples
mean_platesnorm_pPIR1WT_exp_rep1 <- platesnorm_pPIR1_exp_rep  %>% filter(exp_rep==1)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pPIR1-mCherry-tPIR1_WT")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_WT = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_WT = mean(median_bio_rep_WT))

# Calculates the mean cq value of pPIR1-mCherry-tPIR1_WT Samples
mean_platesnorm_pPIR1WT_exp_rep2 <- platesnorm_pPIR1_exp_rep  %>% filter(exp_rep==2)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pPIR1-mCherry-tPIR1_WT")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_WT = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_WT = mean(median_bio_rep_WT))

# Calculates the delta delta cq (subtracts mean(pPIR1WT$delta_cq) from every delta_cq))
platesnorm_WT_pPIR1_exp_rep1 <- platesnorm_pPIR1_exp_rep %>% filter(exp_rep==1)%>%
    mutate(delta_cq.WT = delta_cq - mean_platesnorm_pPIR1WT_exp_rep1$mean_WT, remove=FALSE)
platesnorm_WT_pPIR1_exp_rep2 <- platesnorm_pPIR1_exp_rep %>% filter(exp_rep==2)%>%
    mutate(delta_cq.WT = delta_cq - mean_platesnorm_pPIR1WT_exp_rep2$mean_WT, remove=FALSE)
```


```{r delta-deltacq_norm_PGK1_exp_rep, echo=TRUE}

# Extracts the pPGK1 strains
platesnorm_pPGK1_exp_rep <- platesnorm %>% filter(promoter %in% "pPGK1", target_id %in% "mCh-7")
  
# Calculates the mean cq value of pPGK1-mCherry-tPIR1_WT Samples
mean_platesnorm_pPGK1WT_exp_rep1 <- platesnorm_pPGK1_exp_rep  %>% filter(exp_rep==1)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pPGK1-mCherry-tPIR1_WT")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_WT = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_WT = mean(median_bio_rep_WT))

# Calculates the mean cq value of pPGK1-mCherry-tPIR1_WT Samples
mean_platesnorm_pPGK1WT_exp_rep2 <- platesnorm_pPGK1_exp_rep  %>% filter(exp_rep==2)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pPGK1-mCherry-tPIR1_WT")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_WT = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_WT = mean(median_bio_rep_WT))

# Calculates the delta delta cq (subtracts mean(pPGK1WT$delta_cq) from every delta_cq))
platesnorm_WT_pPGK1_exp_rep1 <- platesnorm_pPGK1_exp_rep %>% filter(exp_rep==1)%>%
    mutate(delta_cq.WT = delta_cq - mean_platesnorm_pPGK1WT_exp_rep1$mean_WT, remove=FALSE)
platesnorm_WT_pPGK1_exp_rep2 <- platesnorm_pPGK1_exp_rep %>% filter(exp_rep==2)%>%
    mutate(delta_cq.WT = delta_cq - mean_platesnorm_pPGK1WT_exp_rep2$mean_WT, remove=FALSE)
```

```{r delta-deltacq_norm_SRO9_exp_rep, echo=TRUE}

# Extracts the SRO9 strains
platesnorm_pSRO9_exp_rep <- platesnorm %>% filter(promoter %in% "pSRO9", target_id %in% "mCh-7")
  
# Calculates the mean cq value of pSRO9-mCherry-tPIR1_WT Samples 
mean_platesnorm_pSRO9WT_exp_rep1 <- platesnorm_pSRO9_exp_rep  %>% filter(exp_rep==1)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pSRO9-mCherry-tPIR1_WT")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_WT = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_WT = mean(median_bio_rep_WT))

# Calculates the mean cq value of pSRO9-mCherry-tPIR1_WT Samples
mean_platesnorm_pSRO9WT_exp_rep2 <- platesnorm_pSRO9_exp_rep  %>% filter(exp_rep==2)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pSRO9-mCherry-tPIR1_WT")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_WT = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_WT = mean(median_bio_rep_WT))

# Calculates the delta delta cq (subtracts mean(pPGK1WT$delta_cq) from every delta_cq))
platesnorm_WT_pSRO9_exp_rep1 <- platesnorm_pSRO9_exp_rep %>% filter(exp_rep==1)%>%
    mutate(delta_cq.WT = delta_cq - mean_platesnorm_pSRO9WT_exp_rep1$mean_WT, remove=FALSE)
platesnorm_WT_pSRO9_exp_rep2 <- platesnorm_pSRO9_exp_rep %>% filter(exp_rep==2)%>%
    mutate(delta_cq.WT = delta_cq - mean_platesnorm_pSRO9WT_exp_rep2$mean_WT, remove=FALSE)
```

## 2) Mean delta delta cq and RNA abundance calculation
In the previous 2 chunks, we have calculated the delta-delta cq value for each techrep. Here, we are summarising that information by calculating the median cq for each biorep and calculating the RNA abundance (2^-Median_cq).  

```{r mean_rna_abund_exp_rep, echo=TRUE}

# Combining the calculated delta-delta cq from the two dataframes then calculating the Median delta-delta cq. RNA abundance is calculated by the following formula 2^(- Median_cq). 
platesnorm_WT_median_exp_rep <- bind_rows(platesnorm_WT_pPIR1_exp_rep1, platesnorm_WT_pPIR1_exp_rep2,
                                            platesnorm_WT_pPGK1_exp_rep1, platesnorm_WT_pPGK1_exp_rep2,
                                            platesnorm_WT_pSRO9_exp_rep1, platesnorm_WT_pSRO9_exp_rep2)%>%
                    group_by(strain, target_id, bio_rep, exp_rep)%>%
                    summarize(delta_deltacq = median(delta_cq.WT, na.rm=TRUE),
                              rel_abund_delta_deltacq = (2^-delta_deltacq), na.rm=FALSE)%>%
                    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
                    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
                    mutate(construct = factor(construct,levels = c("modG","modF", "modE","modD","modC","modA","modB","WT")),remove = FALSE)%>%
                    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
                    unite(UTR3,terminator,construct,sep="_", remove=FALSE) %>%
                    mutate(pro_mCh = factor(pro_mCh,levels = c("pPIR1-mCherry","pPGK1-mCherry", "pSRO9-mCherry"))) %>% 
                    mutate(UTR3 = factor(UTR3,levels=c("tPIR1_modG", "tPIR1_modF", "tPIR1_modE", "tPIR1_modD","tPIR1_modC","tPIR1_modA","tPIR1_modB", "tPIR1_WT")))
```

### 2) Plotting the results in a scatter plot

```{r plot_delta_deltacq_exp_rep_1,fig.height=5,fig.width=9, echo=FALSE}

normalised_plot_exp_rep <- ggplot(data = platesnorm_WT_median_exp_rep) +
  RNA_relative_abundance_figure_options +
  scale_colour_manual(values=PIR1_colour_scheme) +
    theme(axis.text.y=element_text(colour=PIR1_colour_scheme)) +
  labs(x ="Fold change in RNA abundance relative to tPIR1-mod0 (log2 scale)", y = "") +
  facet_wrap(~pro_mCh,ncol = 3)
          

ggsave(here("results_chapter/figures/tPIR1_design_and_qpcr.png"),
       plot_grid(ggdraw() + 
                   draw_image(here("raw_data_analysis/figures/terminator_construct_designs/PIR1_motif_insertion_construct_design.svg")),
                 normalised_plot_exp_rep, 
                 ncol = 1, 
                 labels = c("A", "B"), 
                 rel_heights =  c(1,0.7)), 
       width = 7, 
       height = 7)
```

```{r plot_delta_deltacq_exp_rep_2,fig.height=9,fig.width=5, echo=FALSE}

normalised_plot1_exp_rep <- ggplot(data = platesnorm_WT_median_exp_rep) +
    RNA_relative_abundance_figure_options +
  scale_colour_manual(values=PIR1_colour_scheme) +
    theme(axis.text.y=element_text(colour=PIR1_colour_scheme)) +
  labs(x ="Fold change in RNA abundance relative to tPIR1-mod0 (log2 scale)", y = "") +
  facet_wrap(~pro_mCh,ncol = 3)

normalised_plot1_exp_rep
```
