---
title: "RT-qPCR of pTSA1-mCh-tTSA1mod (2 reps), pPGK1-mCh-tTSA1mod (2 reps) and pSRO9-mCh-tTSA1mod (2 reps)"
author: "Jamie Auxillos"
date: "06/05/2020"
output:
  html_document:
    toc: yes
    #toc_depth: 2
---

```{r setup,warning=FALSE,message=FALSE,echo=FALSE}

## knitr options for report generation
knitr::opts_chunk$set(warning=FALSE,message=FALSE,echo=TRUE,cache=FALSE)


library(dplyr)
library(tibble)
library(stringr)
library(readr)
library(cowplot)
library(tidyqpcr)
library(here)

 # set default theme for graphics
source(here("raw_data_analysis/code/shared_figure_formatting.R"))

```

```{r label_plates, echo=FALSE}

# list Targets (Primer sets)
target_id <- c("mCh-7", "URA3-ORF", "TSA1-ORF", "PGK1-ORF")
target_id_values <-factor(rep(target_id,levels=target_id))


# labelling of SampleID attributes
strain_levels_TSA1 <- c("pTSA1-mCherry-tTSA1_WT","pTSA1-mCherry-tTSA1_mod0","pTSA1-mCherry-tTSA1_modA","pTSA1-mCherry-tTSA1_modB","pTSA1-mCherry-tTSA1_modC", "pTSA1-mCherry-tTSA1_modD", "pTSA1-mCherry-tTSA1_modE","POT1-ccdB")
strain_values_TSA1 <- factor(rep(strain_levels_TSA1,each=3),levels=strain_levels_TSA1)
strain_levels_PGK1 <- c("pPGK1-mCherry-tTSA1_WT","pPGK1-mCherry-tTSA1_mod0","pPGK1-mCherry-tTSA1_modA","pPGK1-mCherry-tTSA1_modB","pPGK1-mCherry-tTSA1_modC", "pPGK1-mCherry-tTSA1_modD", "pPGK1-mCherry-tTSA1_modE","POT1-ccdB")
strain_values_PGK1 <- factor(rep(strain_levels_PGK1,each=3),levels=strain_levels_PGK1)
strain_levels_SRO9 <- c("pSRO9-mCherry-tTSA1_WT","pSRO9-mCherry-tTSA1_mod0","pSRO9-mCherry-tTSA1_modA","pSRO9-mCherry-tTSA1_modB","pSRO9-mCherry-tTSA1_modC", "pSRO9-mCherry-tTSA1_modD", "pSRO9-mCherry-tTSA1_modE","POT1-ccdB")
strain_values_SRO9 <- factor(rep(strain_levels_SRO9,each=3),levels=strain_levels_SRO9)
bio_rep_levels_A <- c("bio_rep1","bio_rep2","bio_rep3")
bio_rep_values_A <- factor(rep(bio_rep_levels_A,times=8),levels=bio_rep_levels_A)
bio_rep_levels_B <- c("bio_rep4","bio_rep5","bio_rep6")
bio_rep_values_B <- factor(rep(bio_rep_levels_B,times=8),levels=bio_rep_levels_B)

# Attribute labels to each row and column for each plate (pTSA1 and pPGK1 plates)
colkey_TSA1_1 <- tibble(well_col=1:24, strain=strain_values_TSA1, bio_rep=bio_rep_values_A) 
colkey_TSA1_2 <- tibble(well_col=1:24, strain=strain_values_TSA1, bio_rep=bio_rep_values_B) 
colkey_PGK1_1 <- tibble(well_col=1:24, strain=strain_values_PGK1, bio_rep=bio_rep_values_A) 
colkey_PGK1_2 <- tibble(well_col=1:24, strain=strain_values_PGK1, bio_rep=bio_rep_values_B) 
colkey_SRO9_1 <- tibble(well_col=1:24, strain=strain_values_SRO9, bio_rep=bio_rep_values_A) 
colkey_SRO9_2 <- tibble(well_col=1:24, strain=strain_values_SRO9, bio_rep=bio_rep_values_B) 
rowkey <- create_rowkey_4_in_16(target_id=target_id_values) 


# Combine all the information into 4x plate plans
plateplan_TSA1_1 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_TSA1_1%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_TSA1_2 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_TSA1_2%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_PGK1_1 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_PGK1_1%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_PGK1_2 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_PGK1_2%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_SRO9_1 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_SRO9_1%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))

plateplan_SRO9_2 <-     
    label_plate_rowcol(create_blank_plate(well_row = LETTERS[1:16],well_col=1:24),
                       rowkey,colkey_SRO9_2%>% 
    unite(sample_id,strain,bio_rep,remove=FALSE)%>%
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE))
```

## Load data and attribute sample information
```{r load_plates,dependson="label_plates",results="show"}
# read my plates


plates_TSA1_1 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pTSA1-tTSA1/JA_20201107-pTSA1-tTSA1mod-n1-ct.txt")) %>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_TSA1_2)

plates_TSA1_2 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pTSA1-tTSA1/JA_20201108-pTSA1-tTSA1mod-n2-ct.txt")) %>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_TSA1_1)

plates_PGK1_1 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tTSA1/JA_20201108-pPGK1-tTSA1mod-n1-ct.txt")) %>%
    mutate(exp_rep = 1, exp_rep=factor(exp_rep)) %>% 
    left_join(plateplan_PGK1_1)

plates_PGK1_2 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tTSA1/JA_20201108-pPGK1-tTSA1mod-n2-ct.txt")) %>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_PGK1_2)

plates_SRO9_1 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tTSA1/JA_20201109-pSRO9-tTSA1mod_n1-ct.txt")) %>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_SRO9_1)

plates_SRO9_2 <- read_lightcycler_1colour_cq(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tTSA1/JA_20201109-pSRO9-tTSA1mod_n2-ct.txt")) %>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep)) %>%
    left_join(plateplan_SRO9_2)


plates <- bind_rows(plates_TSA1_1,plates_TSA1_2,plates_PGK1_1,plates_PGK1_2, plates_SRO9_1, plates_SRO9_2) 

summary(plates)
```


### Plotting unnormalised data for pTSA1 strains
Here we are plotting the unnormalised qPCR data for the constructs containing pTSA1 driving the expression of mCherry. All -RT samples appear to have high cq values and all +RT samples have cq values below 20 cycles. There is however a clear difference in the cq values between experimental replicates (Experimental replicate 1 -Biorep 1, 2, 3)(Experimental replicate 2 -Biorep 4, 5, 6). Experimental replicate 2 seems to have lower cq values across all primers and strains than experimental replicate 1.  

```{r unnorm_pTSA1,fig.height=5,fig.width=10, echo=FALSE}
plates_TSA1 <- plates%>% filter(strain %in% c("pTSA1-mCherry-tTSA1_WT","pTSA1-mCherry-tTSA1_mod0","pTSA1-mCherry-tTSA1_modA","pTSA1-mCherry-tTSA1_modB",
                                           "pTSA1-mCherry-tTSA1_modC","pTSA1-mCherry-tTSA1_modD","pTSA1-mCherry-tTSA1_modE"))%>%
  mutate(strain = factor(strain,levels=c("pTSA1-mCherry-tTSA1_WT","pTSA1-mCherry-tTSA1_mod0","pTSA1-mCherry-tTSA1_modA","pTSA1-mCherry-tTSA1_modB",
                                           "pTSA1-mCherry-tTSA1_modC","pTSA1-mCherry-tTSA1_modD","pTSA1-mCherry-tTSA1_modE")))

ggplot(data=plates_TSA1, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.6, alpha=0.7) +
    labs(y="Quantification cycle (cq)",x="Primers") +
    scale_shape_manual(values=c(16,23)) +
    scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
    facet_wrap(~strain,ncol=4) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3),
          legend.position=c(0.85,0.01))
```


### Plotting unnormalised data for pPGK1 strains
Here we are plotting the unnormalised qPCR data for the constructs containing pTSA1 driving the expression of mCherry. All -RT samples appear to have high cq values and all +RT samples have cq values below 20 cycles. There is little difference between the experimental replicates.

```{r unnorm_pPGK1,fig.height=5,fig.width=10, echo=FALSE}

plates_PGK1 <- plates%>% filter(strain %in% c("pPGK1-mCherry-tTSA1_WT","pPGK1-mCherry-tTSA1_mod0","pPGK1-mCherry-tTSA1_modA","pPGK1-mCherry-tTSA1_modB",
                                           "pPGK1-mCherry-tTSA1_modC","pPGK1-mCherry-tTSA1_modD","pPGK1-mCherry-tTSA1_modE"))%>%
  mutate(strain = factor(strain,levels=c("pPGK1-mCherry-tTSA1_WT","pPGK1-mCherry-tTSA1_mod0","pPGK1-mCherry-tTSA1_modA","pPGK1-mCherry-tTSA1_modB",
                                           "pPGK1-mCherry-tTSA1_modC","pPGK1-mCherry-tTSA1_modD","pPGK1-mCherry-tTSA1_modE")))
ggplot(data=plates_PGK1, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.6, alpha=0.7) +
    labs(y="Quantification cycle (cq)",x="Primers") +
    scale_shape_manual(values=c(16,23)) +
    scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
    facet_wrap(~strain,ncol=4) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3),
          legend.position=c(0.85,0.01))
```

### Plotting unnormalised data for pSRO9 strains
Here we are plotting the unnormalised qPCR data for the constructs containing pTSA1 driving the expression of mCherry. All -RT samples appear to have high cq values and all +RT samples have cq values below 20 cycles. There is little difference between the experimental replicates.
```{r unnorm_pSRO9,fig.height=5,fig.width=10, echo=FALSE}

plates_SRO9 <- plates%>% filter(strain %in% c("pSRO9-mCherry-tTSA1_WT","pSRO9-mCherry-tTSA1_mod0","pSRO9-mCherry-tTSA1_modA","pSRO9-mCherry-tTSA1_modB",
                                           "pSRO9-mCherry-tTSA1_modC","pSRO9-mCherry-tTSA1_modD","pSRO9-mCherry-tTSA1_modE"))%>%
  mutate(strain = factor(strain,levels=c("pSRO9-mCherry-tTSA1_WT","pSRO9-mCherry-tTSA1_mod0","pSRO9-mCherry-tTSA1_modA","pSRO9-mCherry-tTSA1_modB",
                                           "pSRO9-mCherry-tTSA1_modC","pSRO9-mCherry-tTSA1_modD","pSRO9-mCherry-tTSA1_modE")))
ggplot(data=plates_SRO9, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.6, alpha=0.7) +
    labs(y="Quantification cycle (cq)",x="Primers") +
    scale_shape_manual(values=c(16,23)) +
    scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
    facet_wrap(~strain,ncol=4) +
    theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3),
          legend.position=c(0.85,0.01))
```


### Plotting unnormalised data for control strains (pTSA1-mCherry-tTSA1_WT and POT1-ccdB)


```{r unnorm_POT_pTSA1, fig.height=3.5,fig.width=5.5, echo=FALSE}

plates_POT_TSA1 <- bind_rows(plates_TSA1_1,plates_TSA1_2)%>% filter(strain %in% c("pTSA1-mCherry-tTSA1_WT","POT1-ccdB"))

ggplot(data=plates_POT_TSA1, aes(x=target_id,y=cq)) +
  geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.5, alpha=0.7) +
  labs(title="Controls for pTSA1 strains - \n Comparing between experimental replicates",y="Quantification cycle (cq)",x="Primers") +
  ylim(0,40)+
  scale_shape_manual(values=c(16,23)) +
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  facet_wrap(~strain,ncol=2)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
```


### Plotting unnormalised data for control strains (pPGK1-mCherry-tTSA1_WT and POT1-ccdB)


```{r unnorm_POT_pPGK1, fig.height=3.5,fig.width=5.5, echo=FALSE}

plates_POT_PGK1 <- bind_rows(plates_PGK1_1,plates_PGK1_2)%>% filter(strain %in% c("pPGK1-mCherry-tTSA1_WT","POT1-ccdB"))

ggplot(data=plates_POT_PGK1, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.5, alpha=0.7) +
  labs(title="Controls for pPGK1 strains - \n Comparing between experimental replicates",y="Quantification cycle (cq)",x="Primers") +
  ylim(0,40)+
  scale_shape_manual(values=c(16,23)) +
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  facet_wrap(~strain,ncol=2)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
```

### Plotting unnormalised data for control strains (pSRO9-mCherry-tTSA1_WT and POT1-ccdB)

```{r unnorm_POT_pSRO9, fig.height=3.5,fig.width=5.5, echo=FALSE}

plates_POT_SRO9 <- bind_rows(plates_SRO9_1,plates_SRO9_2)%>% filter(strain %in% c("pSRO9-mCherry-tTSA1_WT","POT1-ccdB"))

ggplot(data=plates_POT_SRO9, aes(x=target_id,y=cq)) +
    geom_point(aes(color=bio_rep, shape=prep_type),
               position=position_dodge(width = 0.85),size=1.5, alpha=0.7) +
  labs(title="Controls for pSRO9 strains - \n Comparing between experimental replicates",y="Quantification cycle (cq)",x="Primers") +
  ylim(0,40)+
  scale_shape_manual(values=c(16,23)) +
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  facet_wrap(~strain,ncol=2)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5),
          axis.title.x=element_text(vjust=-2),
          axis.title.y=element_text(vjust=3))
```

## Loading amplification and melt curve pTSA1 data

```{r load_amp-melt_TSA1,dependson="label_plates", results="show"}

platecurve_TSA1_1 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pTSA1-tTSA1/JA_20201107-pTSA1-tTSA1mod-n1.txt"),skip=2,
                      col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_TSA1_1)%>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep))

platecurve_TSA1_2 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pTSA1-tTSA1/JA_20201108-pTSA1-tTSA1mod-n2.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_TSA1_2)%>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep))

platecurve_TSA1 <- bind_rows(platecurve_TSA1_1, platecurve_TSA1_2)

platesamp_TSA1  <- platecurve_TSA1 %>% filter(program_no == 2)
platesmelt_TSA1 <- platecurve_TSA1 %>% filter(program_no != 2) %>% getdRdTall() %>% filter(temperature >= 61)  
```

### Amplification Curves of POT1-ccdB strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, TSA1-ORF, URA3-ORF) (pTSA1)

```{r plotamp_Rep_TSA1,fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesamp_TSA1 %>% filter(strain==c("POT1-ccdB")),
       aes(x=cycle,y=fluor_signal,linetype=prep_type,group=well, color=exp_rep)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line() +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    expand_limits(y=0) + 
    labs(title="Amplification curves of POT1-ccdB strains (bio_rep 1-6) in pTSA1 experiments") + panel_border()

ggplot(data=platesamp_TSA1 %>% filter(strain==c("pTSA1-mCherry-tTSA1_WT")),
       aes(x=cycle,y=fluor_signal,linetype=prep_type,group=well, color=exp_rep)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line() +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    expand_limits(y=0) + 
    labs(title="Amplification curves of pTSA1-mCherry-tTSA1_WT strains (bio_rep 1-6) in pTSA1 experiments") + panel_border()
```



### Melt Curves of POT1-ccdB and pTSA1-mCherry-tTSA1_WT strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, TSA1-ORF, URA3-ORF) (pTSA1)
  

```{r plotmelt_Rep_TSA1,dependson="load_amp",fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesmelt_TSA1 %>% filter(strain==c("POT1-ccdB")),
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of POT1-ccdB strains (bio_rep 1-6) in pTSA1 experiments") + panel_border()

ggplot(data=platesmelt_TSA1 %>% filter(strain==c("pTSA1-mCherry-tTSA1_WT")),
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of pTSA1-mCherry-tTSA1_WT strains (bio_rep 1-6) in pTSA1 experiments") + panel_border()
```


### Loading amplification and melt curve pPGK1 data

```{r load_amp-melt_P1,dependson="label_plates"}

platecurve_PGK1_1 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tTSA1/JA_20201108-pPGK1-tTSA1mod-n1.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_PGK1_1)%>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep))

platecurve_PGK1_2 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pPGK1-tTSA1/JA_20201108-pPGK1-tTSA1mod-n2.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_PGK1_2)%>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep))

platecurve_PGK1 <- bind_rows(platecurve_PGK1_1, platecurve_PGK1_2)

platesamp_PGK1  <- platecurve_PGK1 %>% filter(program_no == 2)
platesmelt_PGK1 <- platecurve_PGK1 %>% filter(program_no != 2) %>% getdRdTall() %>% filter(temperature >= 61)  
```

### Amplification Curves of POT1-ccdB strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, TSA1-ORF, URA3-ORF) (pPGK1)

```{r plotamp_Rep_PGK1,dependson="load_amp",fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesamp_PGK1 %>% filter(strain==c("POT1-ccdB")),
       aes(x=cycle,y=fluor_signal,linetype=prep_type,group=well, color=exp_rep)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line() +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    expand_limits(y=0) + 
    labs(title="Amplification curves of POT1-ccdB strains (bio_rep 1-6) in pPGK1 experiments") + panel_border()
```

### Melt Curves of POT1-ccdB and pPGK1-mCherry-tTSA1_WT strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, TSA1-ORF, URA3-ORF) (pPGK1)

```{r plotmelt_Rep_PGK1,fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesmelt_PGK1 %>% filter(strain==c("POT1-ccdB")),
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of POT1-ccdB strains (bio_rep 1-6) in pPGK1 experiments") + panel_border()

ggplot(data=platesmelt_PGK1 %>% filter(strain==c("pPGK1-mCherry-tTSA1_WT")),
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of pPGK1-mCherry-tTSA1_WT strains (bio_rep 1-6) in pPGK1 experiments") + panel_border()
```

### Loading amplification and melt curve pSRO9 data

```{r load_amp-melt_SRO9,dependson="label_plates"}

platecurve_SRO9_1 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tTSA1/JA_20201109-pSRO9-tTSA1mod_n1.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_SRO9_1)%>%
    mutate(exp_rep = 1,exp_rep=factor(exp_rep))

platecurve_SRO9_2 <- read_tsv(here("raw_data_analysis/data/raw_qpcr/motif_context_dependence/pSRO9-tTSA1/JA_20201109-pSRO9-tTSA1mod_n2.txt"),skip=2,
                           col_names=c("well","SID","program_no","segment","cycle","time","temperature","fluor_raw")) %>%
    debaseline() %>%left_join(plateplan_SRO9_2)%>%
    mutate(exp_rep = 2,exp_rep=factor(exp_rep))

platecurve_SRO9 <- bind_rows(platecurve_SRO9_1, platecurve_SRO9_2)

platesamp_SRO9  <- platecurve_SRO9 %>% filter(program_no == 2)
platesmelt_SRO9 <- platecurve_SRO9 %>% filter(program_no != 2) %>% getdRdTall() %>% filter(temperature >= 61)  
```

### Amplification Curves of POT1-ccdB strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, TSA1-ORF, URA3-ORF) (pSRO9)


```{r plotamp_Rep_SRO9,dependson="load_amp",fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesamp_SRO9 %>% filter(strain==c("POT1-ccdB")),
       aes(x=cycle,y=fluor_signal,linetype=prep_type,group=well, color=exp_rep)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line() +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    expand_limits(y=0) + 
    labs(title="Amplification curves of POT1-ccdB strains (bio_rep 1-6) in pSRO9 experiments") + panel_border()
```

### Melt Curves of POT1-ccdB and pPGK1-mCherry-tTSA1_WT strains (all bioreps) against all 4 target_ids (mCh-7, PGK1-ORF, TSA1-ORF, URA3-ORF) (SRO9)

```{r plotmelt_Rep_SRO9,fig.width=8,fig.height=6, echo=FALSE}
ggplot(data=platesmelt_SRO9 %>% filter(strain==c("POT1-ccdB")),
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of POT1-ccdB strains (bio_rep 1-6) in pSRO9 experiments") + panel_border()

ggplot(data=platesmelt_SRO9 %>% filter(strain==c("pSRO9-mCherry-tTSA1_WT")),
       aes(x=temperature,y=dRdT,group=well)) + 
    facet_grid(bio_rep~target_id) + 
    geom_line(aes(color=exp_rep, linetype=prep_type)) +
    scale_linetype_manual(values=c("-RT"="dashed","+RT"="solid")) +
    scale_x_continuous(breaks=seq(60,100,10),minor_breaks=seq(60,100,5)) + 
    labs(title="Melt curves of pSRO9-mCherry-tTSA1_WT strains (bio_rep 1-6) in pSRO9 experiments") + panel_border()
```


## 1) Delta cq calculation
norm_function changed to **mean**

```{r deltacq_norm, echo=TRUE, fig.height=6, fig.width=9}

# platesnorm first excludes pPGK1-mCherry-tTSA1_WT_bio_rep2, and all POT1-ccdB strains, then takes all +RT results, then normalises the cq values by the median of normtarget_ids, then filters cq values for mCh-7
platesnorm <- plates  %>% 
        filter(!strain %in% c("POT1-ccdB"), prep_type=="+RT") %>%    
        calculate_deltacq_bysampleid(ref_target_ids = c("PGK1-ORF", "RPS3-ORF", "URA3-ORF"), norm_function = mean)

platesnorm_summarise <- platesnorm%>%
                    group_by(strain, target_id, bio_rep, exp_rep)%>%
                    summarize(deltacq = median(delta_cq, na.rm=TRUE),
                              rel_abund_deltacq = (2^-deltacq), na.rm=FALSE)%>% 
    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
    unite(UTR3,terminator,construct,sep="_", remove=FALSE)



ggplot(platesnorm, aes(UTR3,delta_cq))+
  geom_point(aes(color=bio_rep),position=position_dodge(width = 0.85),size=1.2, alpha=0.7)+
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  labs(y="delta cq", x="3'UTR-terminators")+
  facet_wrap(~target_id)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5))+
  geom_hline(yintercept = 0, color = "black", linetype= 'dotted', size=1) 
```

### 1) Exporting analysed delta cq values data
```{r deltacq_normdata, echo=TRUE}
#exporting platesnorm dataframe
write.csv(platesnorm_summarise,here("raw_data_analysis/data/norm_qpcr/motif_context_dependence/pTSA1_pPGK1_pSRO9_tTSA1_deltacq_platesnorm_summarise.csv"), row.names = FALSE)

```


```{r plot_deltacq, echo=TRUE, fig.height=5, fig.width=9}

ggplot(platesnorm %>% filter(target_id %in% 'mCh-7'), aes(UTR3,delta_cq))+
  geom_point(aes(color=exp_rep),position=position_dodge(width = 0.85),size=1.2, alpha=1)+
  scale_colour_hue(h = c(90, 360)+20,l=60,c=60)+
  labs(y="delta cq", x="3'UTR-terminators")+
  facet_wrap(~promoter)+
  theme(axis.text.x=element_text(angle=90,vjust=0.5))+
  geom_hline(yintercept = 0, color = "black", linetype= 'dotted', size=1) 
```

## 2) Delta-delta cq and RNA abundance calculation (one mod0 mean per exp_rep per promoter)
```{r delta-deltacq_norm_TSA1_exp_rep, echo=TRUE}

# Extracts the pTSA1 strains
platesnorm_pTSA1_exp_rep <- platesnorm %>% filter(promoter %in% "pTSA1", target_id %in% "mCh-7")
  
# Calculates the mean cq value of pTSA1-mCherry-tTSA1_mod0 Samples (-0.1133333)
mean_platesnorm_pTSA1mod0_exp_rep1 <- platesnorm_pTSA1_exp_rep  %>% filter(exp_rep==1)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pTSA1-mCherry-tTSA1_mod0")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_mod0 = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_mod0 = mean(median_bio_rep_mod0))

# Calculates the mean cq value of pTSA1-mCherry-tTSA1_mod0 Samples (-1.236667)
mean_platesnorm_pTSA1mod0_exp_rep2 <- platesnorm_pTSA1_exp_rep  %>% filter(exp_rep==2)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pTSA1-mCherry-tTSA1_mod0")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_mod0 = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_mod0 = mean(median_bio_rep_mod0))

# Calculates the delta delta cq (subtracts mean(pTSA1mod0$delta_cq) from every delta_cq))
platesnorm_mod0_pTSA1_exp_rep1 <- platesnorm_pTSA1_exp_rep %>% filter(exp_rep==1)%>%
    mutate(delta_cq.mod0 = delta_cq - mean_platesnorm_pTSA1mod0_exp_rep1$mean_mod0, remove=FALSE)
platesnorm_mod0_pTSA1_exp_rep2 <- platesnorm_pTSA1_exp_rep %>% filter(exp_rep==2)%>%
    mutate(delta_cq.mod0 = delta_cq - mean_platesnorm_pTSA1mod0_exp_rep2$mean_mod0, remove=FALSE)
```


```{r delta-deltacq_norm_PGK1_exp_rep, echo=TRUE}

# Extracts the pPGK1 strains
platesnorm_pPGK1_exp_rep <- platesnorm %>% filter(promoter %in% "pPGK1", target_id %in% "mCh-7")
  
# Calculates the mean cq value of pPGK1-mCherry-tTSA1_mod0 Samples (-2.158333)
mean_platesnorm_pPGK1mod0_exp_rep1 <- platesnorm_pPGK1_exp_rep  %>% filter(exp_rep==1)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pPGK1-mCherry-tTSA1_mod0")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_mod0 = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_mod0 = mean(median_bio_rep_mod0))

# Calculates the mean cq value of pPGK1-mCherry-tTSA1_mod0 Samples (-2.136667)
mean_platesnorm_pPGK1mod0_exp_rep2 <- platesnorm_pPGK1_exp_rep  %>% filter(exp_rep==2)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pPGK1-mCherry-tTSA1_mod0")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_mod0 = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_mod0 = mean(median_bio_rep_mod0))

# Calculates the delta delta cq (subtracts mean(pPGK1mod0$delta_cq) from every delta_cq))
platesnorm_mod0_pPGK1_exp_rep1 <- platesnorm_pPGK1_exp_rep %>% filter(exp_rep==1)%>%
    mutate(delta_cq.mod0 = delta_cq - mean_platesnorm_pPGK1mod0_exp_rep1$mean_mod0, remove=FALSE)
platesnorm_mod0_pPGK1_exp_rep2 <- platesnorm_pPGK1_exp_rep %>% filter(exp_rep==2)%>%
    mutate(delta_cq.mod0 = delta_cq - mean_platesnorm_pPGK1mod0_exp_rep2$mean_mod0, remove=FALSE)
```

```{r delta-deltacq_norm_SRO9_exp_rep, echo=TRUE}

# Extracts the SRO9 strains
platesnorm_pSRO9_exp_rep <- platesnorm %>% filter(promoter %in% "pSRO9", target_id %in% "mCh-7")
  
# Calculates the mean cq value of pSRO9-mCherry-tTSA1_mod0 Samples (4.245926)
mean_platesnorm_pSRO9mod0_exp_rep1 <- platesnorm_pSRO9_exp_rep  %>% filter(exp_rep==1)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pSRO9-mCherry-tTSA1_mod0")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_mod0 = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_mod0 = mean(median_bio_rep_mod0))

# Calculates the mean cq value of pSRO9-mCherry-tTSA1_mod0 Samples (3.887037)
mean_platesnorm_pSRO9mod0_exp_rep2 <- platesnorm_pSRO9_exp_rep  %>% filter(exp_rep==2)%>%
    select(c("strain","delta_cq", "bio_rep", "tech_rep")) %>% 
    filter(strain %in% c("pSRO9-mCherry-tTSA1_mod0")) %>%
    group_by(strain, bio_rep)%>%
    summarize(median_bio_rep_mod0 = median(delta_cq)) %>%
    ungroup()%>%
    group_by(strain)%>%
    summarize(mean_mod0 = mean(median_bio_rep_mod0))

# Calculates the delta delta cq (subtracts mean(pPGK1mod0$delta_cq) from every delta_cq))
platesnorm_mod0_pSRO9_exp_rep1 <- platesnorm_pSRO9_exp_rep %>% filter(exp_rep==1)%>%
    mutate(delta_cq.mod0 = delta_cq - mean_platesnorm_pSRO9mod0_exp_rep1$mean_mod0, remove=FALSE)
platesnorm_mod0_pSRO9_exp_rep2 <- platesnorm_pSRO9_exp_rep %>% filter(exp_rep==2)%>%
    mutate(delta_cq.mod0 = delta_cq - mean_platesnorm_pSRO9mod0_exp_rep2$mean_mod0, remove=FALSE)
```

## 2) Mean delta delta cq and RNA abundance calculation
In the previous 2 chunks, we have calculated the delta-delta cq value for each techrep. Here, we are summarising that information by calculating the median cq for each biorep and calculating the RNA abundance (2^-Median_cq).  

```{r mean_rna_abund_exp_rep, echo=TRUE}

# Combining the calculated delta-delta cq from the two dataframes then calculating the Median delta-delta cq. RNA abundance is calculated by the following formula 2^(- Median_cq). 
platesnorm_mod0_median_exp_rep <- bind_rows(platesnorm_mod0_pTSA1_exp_rep1, platesnorm_mod0_pTSA1_exp_rep2,
                                            platesnorm_mod0_pPGK1_exp_rep1, platesnorm_mod0_pPGK1_exp_rep2,
                                            platesnorm_mod0_pSRO9_exp_rep1, platesnorm_mod0_pSRO9_exp_rep2)%>%
                    group_by(strain, target_id, bio_rep, exp_rep)%>%
                    summarize(delta_deltacq = median(delta_cq.mod0, na.rm=TRUE),
                              rel_abund_delta_deltacq = (2^-delta_deltacq), na.rm=FALSE)%>%
                    separate(strain, remove = FALSE,sep="-",into=c("promoter","mCherry","terminator")) %>%
                    separate(terminator, remove = FALSE, sep="_",into=c("terminator","construct")) %>%
                    mutate(construct = factor(construct,levels = c("modE","modD","modC","modB","modA","mod0","WT")),remove = FALSE)%>%
                    unite(pro_mCh,promoter,mCherry,sep="-", remove=FALSE)%>%
                    unite(UTR3,terminator,construct,sep="_", remove=FALSE) %>%
                    mutate(pro_mCh = factor(pro_mCh,levels = c("pTSA1-mCherry","pPGK1-mCherry", "pSRO9-mCherry"))) %>% 
                    mutate(UTR3 = factor(UTR3,levels=c("tTSA1_WT", "tTSA1_modC","tTSA1_modE","tTSA1_modD","tTSA1_modA",
                                           "tTSA1_modB","tTSA1_mod0")))

write_csv(platesnorm_mod0_median_exp_rep, here("raw_data_analysis/data/norm_qpcr/motif_context_dependence/pTSA1_pPGK1_pSRO9_tTSA1_deltadeltacq_platesnorm_summarise.csv"))

```

### 2) Plotting the results in a scatter plot

```{r plot_delta_deltacq_exp_rep_1,fig.height=5,fig.width=9, echo=FALSE}

normalised_plot_exp_rep <- ggplot(data = platesnorm_mod0_median_exp_rep)+
  RNA_relative_abundance_figure_options +
  scale_colour_manual(values=RPS3_TSA1_colour_scheme) +
   theme(axis.text.y=element_text(colour=RPS3_TSA1_colour_scheme)) +
  labs(x ="Fold change in RNA abundance relative to tTSA1-mod0 (log2 scale)", y = "") +
  facet_wrap(~pro_mCh,ncol = 3)

normalised_plot_exp_rep
```

```{r plot_delta_deltacq_exp_rep_2,fig.height=9,fig.width=5, echo=FALSE}

normalised_plot1_exp_rep <- ggplot(data = platesnorm_mod0_median_exp_rep %>% 
    filter(!strain %in% c("pSRO9-mCherry-tTSA1_WT", "pPGK1-mCherry-tTSA1_WT","pTSA1-mCherry-tTSA1_WT")))+
  RNA_relative_abundance_figure_options +
  scale_colour_manual(values=RPS3_TSA1_colour_scheme) +
   theme(axis.text.y=element_text(colour=RPS3_TSA1_colour_scheme)) +
  labs(x ="Fold change in RNA abundance relative to tTSA1-mod0 (log2 scale)", y = "") +
  facet_wrap(~pro_mCh,ncol = 3)

normalised_plot1_exp_rep
```
