---
title: "qPCR linear models"
output: html_notebook
---

```{r}
library(broom)
library(tidyverse)
library(here)
library(cowplot)
library(ggpattern)
library(latex2exp)

construct_to_motif_dictionary_PIR1 <- tibble(construct = 
                                                    rep(c("WT", "modA", "modB", "modC", "modD", "modE", "modF", "modG"), each = 3), 
                                                  motif = rep(c("ATATTC","TGTACAATA","TTTCATTTC"), times = 8), 
                                                   count = c(1,1,3,
                                                             1,0,3,
                                                             0,1,3,
                                                             1,1,2,
                                                             1,1,2,
                                                             1,1,0,
                                                             1,0,0,
                                                             0,1,0))

construct_to_motif_dictionary_TSA1_RPS3 <- tibble (construct = 
                                                rep(c("WT", "mod0", "modA", "modB", "modC", "modD", "modE"), times = 4), 
                                              motif = 
                                                rep(c("ATATTC", "GTATACCTA","TGTACAATA","TTTCATTTC"), each = 7), 
                                              count = c(0,0,0,2,0,0,0,
                                                        0,0,0,0,2,0,0,
                                                        0,0,1,0,0,0,1,
                                                        0,0,0,0,0,2,2))

pRPS3_pPGK1_pSRO9_tRPS3_norm_mean <- read_csv(here("./raw_data_analysis/data/norm_qpcr/motif_context_dependence/pRPS3_pPGK1_pSRO9_tRPS3_deltacq_platesnorm_summarise.csv")) %>%
  inner_join(construct_to_motif_dictionary_TSA1_RPS3) %>%
  filter( target_id == "mCh-7") %>%
  spread(key= motif, value = count) %>%
  mutate(native = (construct == "WT"), delta_cq = deltacq)

pTSA1_pPGK1_pSRO9_tTSA1_norm_mean <- read_csv(here("./raw_data_analysis/data/norm_qpcr/motif_context_dependence/pTSA1_pPGK1_pSRO9_tTSA1_deltacq_platesnorm_summarise.csv")) %>%
  inner_join(construct_to_motif_dictionary_TSA1_RPS3) %>%
  filter(target_id == "mCh-7") %>%
  spread(key= motif, value = count) %>%
  mutate(native = (construct == "WT"), delta_cq = deltacq)

pPIR1_pPGK1_pSRO9_tPIR1_norm_mean <- read_csv(here("./raw_data_analysis/data/norm_qpcr/motif_context_dependence/pPIR1_pPGK1_pSRO9_tPIR1_deltacq_platesnorm_summarise.csv")) %>%
  inner_join(construct_to_motif_dictionary_PIR1) %>%
  filter(target_id == "mCh-7") %>%
  spread(key= motif, value = count) %>%
  mutate(native = (construct == "WT"), delta_cq = deltacq)

collated_norm_mean_data <- pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>%
  bind_rows(pTSA1_pPGK1_pSRO9_tTSA1_norm_mean) %>%
  bind_rows(pPIR1_pPGK1_pSRO9_tPIR1_norm_mean)

```

```{r run_linear_models, echo=FALSE, include=FALSE}

pTSA1_tTSA1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pTSA1_pPGK1_pSRO9_tTSA1_norm_mean %>% filter(promoter == "pTSA1"))

pPGK1_tTSA1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pTSA1_pPGK1_pSRO9_tTSA1_norm_mean %>% filter(promoter == "pPGK1"))

pSRO9_tTSA1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pTSA1_pPGK1_pSRO9_tTSA1_norm_mean %>% filter(promoter == "pSRO9"))

pRPS3_tRPS3_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>% filter(promoter == "pRPS3"))

pPGK1_tRPS3_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>% filter(promoter == "pPGK1"))

pSRO9_tRPS3_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>% filter(promoter == "pSRO9"))

pPIR1_tPIR1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pPIR1_pPGK1_pSRO9_tPIR1_norm_mean %>% filter(promoter == "pPIR1"))

pPGK1_tPIR1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pPIR1_pPGK1_pSRO9_tPIR1_norm_mean %>% filter(promoter == "pPGK1"))

pSRO9_tPIR1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + native, pPIR1_pPGK1_pSRO9_tPIR1_norm_mean %>% filter(promoter == "pSRO9"))

collated_norm_mean_lm_comparison <- step(lm(delta_cq ~ exp_rep + exp_rep:promoter + exp_rep:terminator + native + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TGTACAATA:TTTCATTTC + terminator + promoter, collated_norm_mean_data), delta_cq ~ exp_rep + exp_rep:promoter + exp_rep:terminator + native + ATATTC + GTATACCTA + TGTACAATA + TTTCATTTC + TTTCATTTC:TGTACAATA + terminator + promoter + terminator:ATATTC + promoter:ATATTC + terminator:GTATACCTA + promoter:GTATACCTA + terminator:TGTACAATA + promoter:TGTACAATA + terminator:TTTCATTTC + promoter:TTTCATTTC + TTTCATTTC:TGTACAATA:terminator + TTTCATTTC:TGTACAATA:promoter + native:terminator + native:promoter)

collated_norm_mean_lm <- lm(delta_cq ~ exp_rep + native + ATATTC + TGTACAATA + TTTCATTTC + GTATACCTA + terminator + promoter + exp_rep:promoter + exp_rep:terminator + TGTACAATA:TTTCATTTC + ATATTC:terminator + TGTACAATA:promoter + ATATTC:promoter + native:promoter + TGTACAATA:terminator + TTTCATTTC:terminator + TTTCATTTC:promoter + TGTACAATA:TTTCATTTC:terminator + TGTACAATA:TTTCATTTC:promoter, collated_norm_mean_data)
```

## Combined linear model results
A greedy algorithm for maximising AIC was used to select for the best combination of terms. First the basic model predicting delta_cq using motif presence, terminator, promoter and exp_rep (with interactive term for exp_rep according to promoter/terminator pairing) was ran giving an adjusted R-squared of 0.7816. Then interactive terms for motif contributions in differing contexts and the linear combination of motifs was added increasing the AIC from  AIC=-197.29 to a maximum of -332.62. The final model after selection is summarised below with an adjusted R-squared of 0.90976.

ATATTC has a strong destabilising effect in RPS3 terminators but not TSA1 terminators, irrespective of promoter but even greater with pRPS3 than pPGK1. TGTACAATA only has a significant effect on its own in TSA terminators. GTATACCTA does not have a significant effect in any context. TTTCATTTC has an significant effect in all context on its own. TTTCATTTC and TGTACAATA have an effect when they appear together in the TSA1 terminators only.
```{r combined_linear_model}
summary(collated_norm_mean_lm)
```

```{r}

model_r_squares <- tibble(
  ter_pro = c("pTSA1_tTSA1", "pPGK1_tTSA1", "pSRO9_tTSA1", "pRPS3_tRPS3", "pPGK1_tRPS3", "pSRO9_tRPS3", "pPIR1_tPIR1", "pPGK1_tPIR1", "pSRO9_tPIR1"),
  R_squared = c(summary(pTSA1_tTSA1_norm_mean_lm)$r.squared,
                summary(pPGK1_tTSA1_norm_mean_lm)$r.squared,
                summary(pSRO9_tTSA1_norm_mean_lm)$r.squared,
                summary(pRPS3_tRPS3_norm_mean_lm)$r.squared,
                summary(pPGK1_tRPS3_norm_mean_lm)$r.squared,
                summary(pSRO9_tRPS3_norm_mean_lm)$r.squared,
                summary(pPIR1_tPIR1_norm_mean_lm)$r.squared,
                summary(pPGK1_tPIR1_norm_mean_lm)$r.squared,
                summary(pSRO9_tPIR1_norm_mean_lm)$r.squared))

model_coefficients <- tidy(pPGK1_tRPS3_norm_mean_lm) %>%
  mutate(ter_pro = "pPGK1_tRPS3", norm_by = "mean") %>%
  bind_rows(tidy(pRPS3_tRPS3_norm_mean_lm) %>%
  mutate(ter_pro = "pRPS3_tRPS3", norm_by = "mean")) %>%
  bind_rows(tidy(pPGK1_tTSA1_norm_mean_lm) %>%
  mutate(ter_pro = "pPGK1_tTSA1", norm_by = "mean")) %>%
  bind_rows(tidy(pTSA1_tTSA1_norm_mean_lm) %>%
  mutate(ter_pro = "pTSA1_tTSA1", norm_by = "mean")) %>%
  bind_rows(tidy(pSRO9_tRPS3_norm_mean_lm) %>%
  mutate(ter_pro = "pSRO9_tRPS3", norm_by = "mean")) %>%
  bind_rows(tidy(pSRO9_tTSA1_norm_mean_lm) %>%
  mutate(ter_pro = "pSRO9_tTSA1", norm_by = "mean"))%>%
  bind_rows(tidy(pSRO9_tPIR1_norm_mean_lm) %>%
  mutate(ter_pro = "pSRO9_tPIR1", norm_by = "mean")) %>%
  bind_rows(tidy(pPIR1_tPIR1_norm_mean_lm) %>%
  mutate(ter_pro = "pPIR1_tPIR1", norm_by = "mean")) %>%
  bind_rows(tidy(pPGK1_tPIR1_norm_mean_lm) %>%
  mutate(ter_pro = "pPGK1_tPIR1", norm_by = "mean"))
  

ggplot(model_coefficients) +
  geom_point(aes(x=term, y=estimate, shape = factor(ter_pro, levels =c("pRPS3_tRPS3", "pTSA1_tTSA1", "pPGK1_tRPS3", "pPGK1_tTSA1")), colour = (p.value < 0.05)), size = 3, position = position_jitter(width = 0.1)) +
  theme_cowplot() +
  theme(axis.text.x = element_text(angle = 90), panel.grid.major.y = element_line(size = 0.1)) +
  labs(x = "", y = "Coefficient", shape = "Promoter/Terminator", colour = "Significant")


motif_term_dictionary <- model_coefficients %>% 
  distinct(term) %>%
  mutate(new_term = c("Construct", "exp_repExpRep2", "ATATTC", "GTATACCTA", "TGTACAATA", "TTTCATTTC", "Wildtype\n vs Mod0", "TGTACAATA\n+\nTTTCATTTC"))

construct_dictionary <- model_coefficients %>% 
  distinct(ter_pro) %>%
  mutate(promoter_type = c("PGK1", "Native", "PGK1", "Native", "SRO9", "SRO9", "SRO9", "Native", "PGK1"))

individual_linear_model_coefficients_points <- ggplot(model_coefficients %>% 
         inner_join(motif_term_dictionary) %>%
         inner_join(construct_dictionary) %>%
         separate(ter_pro, into=c("promoter","terminator"), sep = "_", remove = FALSE) %>%
         filter(!term %in% c("(Intercept)", "nativeTRUE", "exp_rep")) %>%
         arrange(term,promoter),
       aes(x=factor(promoter_type, levels = c("Native","PGK1","SRO9")), y=-estimate, colour = terminator),
    position = position_dodge2()) +
  geom_point(
    size = 5
  ) +
  geom_errorbar(aes(ymax = -estimate + std.error, ymin = -estimate - std.error),
                size =1) +
  labs(y = TeX("Log_2 Fold Change in RNA Abundance"), 
       colour = "Terminator") + 
  theme_bw(base_size = 16) +
  theme(strip.placement = "outside", 
        strip.background = element_rect(fill = "white", colour = "white"), 
        axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0, units = "cm"),
        panel.grid.major.x = element_blank(),
        axis.line.x = element_blank()) +
  scale_colour_manual(breaks = c("tTSA1", "tRPS3", "tPIR1"), values = c("#56B4E9", "#D55E00", "#000001")) +
  facet_grid(~factor(new_term, c("Construct", "Wildtype\n vs Mod0", "ATATTC", "GTATACCTA", "TGTACAATA", "TTTCATTTC", "TGTACAATA\n+\nTTTCATTTC")), 
             scales = "free_x", 
             space = "free_x", 
             switch = "x") +
  geom_hline(yintercept = 0)

ggsave(here("results_chapter/figures/motif_multi_context_coefficient.png"), individual_linear_model_coefficients_points, width = 12, height = 9)

save(individual_linear_model_coefficients_points, file = here("raw_data_analysis/data/motif_multi_context_coefficient.Rdata"))

```