---
title: "qPCR linear models"
output: html_document
---

```{r}
library(broom)
library(dplyr)
library(tibble)
library(stringr)
library(readr)
library(tidyr)
library(here)
library(cowplot)
library(latex2exp)
 # set default theme for graphics
source(here("raw_data_analysis/code/shared_figure_formatting.R"))

construct_to_motif_dictionary_PIR1 <- tibble(construct = 
                                                    rep(c("WT", "modA", "modB", "modC", "modD", "modE", "modF", "modG"), each = 3), 
                                                  motif = rep(c("ATATTC","UGUAHMNUA","HWNCAUUWY"), times = 8), 
                                                   count = c(1,1,3,
                                                             1,0,3,
                                                             0,1,3,
                                                             1,1,2,
                                                             1,1,2,
                                                             1,1,0,
                                                             1,0,0,
                                                             0,1,0))

construct_to_motif_dictionary_TSA1_RPS3 <- tibble (construct = 
                                                rep(c("WT", "mod0", "modA", "modB", "modC", "modD", "modE"), times = 4), 
                                              motif = 
                                                rep(c("ATATTC", "GTATACCTA","UGUAHMNUA","HWNCAUUWY"), each = 7), 
                                              count = c(0,0,0,2,0,0,0,
                                                        0,0,0,0,2,0,0,
                                                        0,0,1,0,0,0,1,
                                                        0,0,0,0,0,2,2))

pRPS3_pPGK1_pSRO9_tRPS3_norm_mean <- read_csv(here("./raw_data_analysis/data/norm_qpcr/motif_context_dependence/pRPS3_pPGK1_pSRO9_tRPS3_deltacq_platesnorm_summarise.csv")) %>%
  inner_join(construct_to_motif_dictionary_TSA1_RPS3) %>%
  filter( target_id == "mCh-7") %>%
  spread(key= motif, value = count) %>%
  mutate(native = (construct == "WT"), delta_cq = deltacq)

pTSA1_pPGK1_pSRO9_tTSA1_norm_mean <- read_csv(here("./raw_data_analysis/data/norm_qpcr/motif_context_dependence/pTSA1_pPGK1_pSRO9_tTSA1_deltacq_platesnorm_summarise.csv")) %>%
  inner_join(construct_to_motif_dictionary_TSA1_RPS3) %>%
  filter(target_id == "mCh-7") %>%
  spread(key= motif, value = count) %>%
  mutate(native = (construct == "WT"), delta_cq = deltacq)

pPIR1_pPGK1_pSRO9_tPIR1_norm_mean <- read_csv(here("./raw_data_analysis/data/norm_qpcr/motif_context_dependence/pPIR1_pPGK1_pSRO9_tPIR1_deltacq_platesnorm_summarise.csv")) %>%
  inner_join(construct_to_motif_dictionary_PIR1) %>%
  filter(target_id == "mCh-7") %>%
  spread(key= motif, value = count) %>%
  mutate(native = (construct == "WT"), delta_cq = deltacq)

collated_norm_mean_data <- pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>%
  bind_rows(pTSA1_pPGK1_pSRO9_tTSA1_norm_mean) %>%
  bind_rows(pPIR1_pPGK1_pSRO9_tPIR1_norm_mean) %>%
  mutate(terminator = factor(terminator, levels = c("tRPS3","tTSA1","tPIR1")))

```

```{r run_linear_models, echo=FALSE, include=FALSE}

pTSA1_tTSA1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pTSA1_pPGK1_pSRO9_tTSA1_norm_mean %>% filter(promoter == "pTSA1"))

pPGK1_tTSA1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pTSA1_pPGK1_pSRO9_tTSA1_norm_mean %>% filter(promoter == "pPGK1"))

pSRO9_tTSA1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pTSA1_pPGK1_pSRO9_tTSA1_norm_mean %>% filter(promoter == "pSRO9"))

pRPS3_tRPS3_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>% filter(promoter == "pRPS3"))

pPGK1_tRPS3_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>% filter(promoter == "pPGK1"))

pSRO9_tRPS3_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pRPS3_pPGK1_pSRO9_tRPS3_norm_mean %>% filter(promoter == "pSRO9"))

pPIR1_tPIR1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pPIR1_pPGK1_pSRO9_tPIR1_norm_mean %>% filter(promoter == "pPIR1"))

pPGK1_tPIR1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pPIR1_pPGK1_pSRO9_tPIR1_norm_mean %>% filter(promoter == "pPGK1"))

pSRO9_tPIR1_norm_mean_lm <- lm(delta_cq ~ exp_rep + ATATTC + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + native, pPIR1_pPGK1_pSRO9_tPIR1_norm_mean %>% filter(promoter == "pSRO9"))

collated_norm_mean_lm_comparison <- step(lm(delta_cq ~ exp_rep + exp_rep:promoter + exp_rep:terminator + native + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + UGUAHMNUA:HWNCAUUWY + terminator + promoter, collated_norm_mean_data), delta_cq ~ exp_rep + exp_rep:promoter + exp_rep:terminator + native + ATATTC + GTATACCTA + UGUAHMNUA + HWNCAUUWY + HWNCAUUWY:UGUAHMNUA + terminator + promoter + terminator:ATATTC + promoter:ATATTC + terminator:GTATACCTA + promoter:GTATACCTA + terminator:UGUAHMNUA + promoter:UGUAHMNUA + terminator:HWNCAUUWY + promoter:HWNCAUUWY + HWNCAUUWY:UGUAHMNUA:terminator + HWNCAUUWY:UGUAHMNUA:promoter + native:terminator + native:promoter)

collated_norm_mean_lm <- lm(delta_cq ~ exp_rep + native + ATATTC + UGUAHMNUA + HWNCAUUWY + GTATACCTA + terminator + promoter + exp_rep:promoter + exp_rep:terminator + UGUAHMNUA:HWNCAUUWY + ATATTC:terminator + UGUAHMNUA:promoter + ATATTC:promoter + native:promoter + UGUAHMNUA:terminator + HWNCAUUWY:terminator + HWNCAUUWY:promoter + UGUAHMNUA:HWNCAUUWY:terminator + UGUAHMNUA:HWNCAUUWY:promoter, collated_norm_mean_data)
```

## Combined linear model results
A greedy algorithm for maximising AIC was used to select for the best combination of terms. First the basic model predicting delta_cq using motif presence, terminator, promoter and exp_rep (with interactive term for exp_rep according to promoter/terminator pairing) was ran giving an adjusted R-squared of 0.7816. Then interactive terms for motif contributions in differing contexts and the linear combination of motifs was added increasing the AIC from  AIC=-197.29 to a maximum of -332.62. The final model after selection is summarised below with an adjusted R-squared of 0.90976.

ATATTC has a strong destabilising effect in RPS3 terminators but not TSA1 terminators, irrespective of promoter but even greater with pRPS3 than pPGK1. UGUAHMNUA only has a significant effect on its own in TSA terminators. GTATACCTA does not have a significant effect in any context. HWNCAUUWY has an significant effect in all context on its own. HWNCAUUWY and UGUAHMNUA have an effect when they appear together in the TSA1 terminators only.
```{r combined_linear_model}
summary(collated_norm_mean_lm)
```

```{r}

model_r_squares <- tibble(
  ter_pro = c("pTSA1_tTSA1", "pPGK1_tTSA1", "pSRO9_tTSA1", "pRPS3_tRPS3", "pPGK1_tRPS3", "pSRO9_tRPS3", "pPIR1_tPIR1", "pPGK1_tPIR1", "pSRO9_tPIR1"),
  R_squared = c(summary(pTSA1_tTSA1_norm_mean_lm)$r.squared,
                summary(pPGK1_tTSA1_norm_mean_lm)$r.squared,
                summary(pSRO9_tTSA1_norm_mean_lm)$r.squared,
                summary(pRPS3_tRPS3_norm_mean_lm)$r.squared,
                summary(pPGK1_tRPS3_norm_mean_lm)$r.squared,
                summary(pSRO9_tRPS3_norm_mean_lm)$r.squared,
                summary(pPIR1_tPIR1_norm_mean_lm)$r.squared,
                summary(pPGK1_tPIR1_norm_mean_lm)$r.squared,
                summary(pSRO9_tPIR1_norm_mean_lm)$r.squared))

model_coefficients <- tidy(pPGK1_tRPS3_norm_mean_lm) %>%
  mutate(ter_pro = "pPGK1_tRPS3", norm_by = "mean") %>%
  bind_rows(tidy(pRPS3_tRPS3_norm_mean_lm) %>%
  mutate(ter_pro = "pRPS3_tRPS3", norm_by = "mean")) %>%
  bind_rows(tidy(pPGK1_tTSA1_norm_mean_lm) %>%
  mutate(ter_pro = "pPGK1_tTSA1", norm_by = "mean")) %>%
  bind_rows(tidy(pTSA1_tTSA1_norm_mean_lm) %>%
  mutate(ter_pro = "pTSA1_tTSA1", norm_by = "mean")) %>%
  bind_rows(tidy(pSRO9_tRPS3_norm_mean_lm) %>%
  mutate(ter_pro = "pSRO9_tRPS3", norm_by = "mean")) %>%
  bind_rows(tidy(pSRO9_tTSA1_norm_mean_lm) %>%
  mutate(ter_pro = "pSRO9_tTSA1", norm_by = "mean"))%>%
  bind_rows(tidy(pSRO9_tPIR1_norm_mean_lm) %>%
  mutate(ter_pro = "pSRO9_tPIR1", norm_by = "mean")) %>%
  bind_rows(tidy(pPIR1_tPIR1_norm_mean_lm) %>%
  mutate(ter_pro = "pPIR1_tPIR1", norm_by = "mean")) %>%
  bind_rows(tidy(pPGK1_tPIR1_norm_mean_lm) %>%
  mutate(ter_pro = "pPGK1_tPIR1", norm_by = "mean"))

# swap sign of PIR1 HWNCAUUWY:UGUAHMNUA to align with direction of change
model_coefficients_PIR1_align <- model_coefficients %>%
  filter(!(term == "UGUAHMNUA:HWNCAUUWY" & str_detect(ter_pro,"tPIR1"))) %>%
  bind_rows( model_coefficients %>%
               filter(term == "UGUAHMNUA:HWNCAUUWY", str_detect(ter_pro,"tPIR1")) %>%
               mutate(estimate = -estimate))

ggplot(model_coefficients_PIR1_align) +
  geom_point(aes(x=term, y=estimate, shape = factor(ter_pro, levels =c("pRPS3_tRPS3", "pTSA1_tTSA1", "pPGK1_tRPS3", "pPGK1_tTSA1")), colour = (p.value < 0.05)), position = position_jitter(width = 0.1)) +
  theme_cowplot() +
  theme(axis.text.x = element_text(angle = 90), panel.grid.major.y = element_line(size = 0.1)) +
  labs(x = "", y = "Coefficient", shape = "Promoter/Terminator", colour = "Significant")


motif_term_dictionary <- model_coefficients_PIR1_align %>% 
  distinct(term) %>%
  mutate(new_term = c("Construct", "exp_repExpRep2", "ATATTC", "GTATACCTA", "TGTAHMNTA", "HWNCATTWY", "Wildtype\n vs Mod0", "TGTAHMNTA\n+\nHWNCATTWY"))

construct_dictionary <- model_coefficients_PIR1_align %>% 
  distinct(ter_pro) %>%
  mutate(promoter_type = c("PGK1", "Native", "PGK1", "Native", "SRO9", "SRO9", "SRO9", "Native", "PGK1"))

individual_lm_coefficients_data <- 
  model_coefficients_PIR1_align %>% 
  inner_join(motif_term_dictionary) %>%
  inner_join(construct_dictionary) %>%
  separate(ter_pro, into=c("promoter","terminator"), sep = "_", remove = FALSE) %>%
  mutate(terminator = factor(terminator, levels = c("tRPS3","tTSA1","tPIR1")),
         promoter_type = factor(promoter_type, levels = c("Native","PGK1","SRO9"))
         ) %>% 
  filter(!term %in% c("(Intercept)", "nativeTRUE", "exp_rep")) %>%
  arrange(term,promoter)

individual_linear_model_coefficients_points <- 
  ggplot(individual_lm_coefficients_data,
       aes(x = promoter_type, 
           y = -estimate, 
           colour = terminator),
    ) +
  geom_pointrange(aes(ymax = -estimate + std.error, ymin = -estimate - std.error),
                position = position_dodge2(width = 0.7),
                fatten = 1) +
  labs(y = expression(atop(log[2]~fold~effect,"on RNA abundance")), 
       colour = "Terminator", tag = "") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5),
        axis.ticks.x = element_blank(),
        strip.text = element_text(size = 8, face = "bold"),
        panel.spacing = unit(0, units = "cm"),
        panel.grid.major.x = element_blank(),
        axis.line.x = element_blank()) +
  scale_colour_manual(values = c("#56B4E9", "#D55E00", "#000001"),
                      guide = guide_legend(override.aes = list(size=0.5))) +
  facet_grid(~factor(new_term, c("Construct", "Wildtype\n vs Mod0", "ATATTC", "GTATACCTA", "TGTAHMNTA", "HWNCATTWY", "TGTAHMNTA\n+\nHWNCATTWY")), 
             scales = "free_x", 
             space = "free_x") +
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.2, colour = "black")  +
  geom_vline(xintercept = c(1.5,2.5), colour = "grey50", size = 0.1)

individual_linear_model_coefficients_points
```

```{r save-results, eval = FALSE}
# replace by saving a separate figure panel???
save(individual_linear_model_coefficients_points, file = here("raw_data_analysis/data/motif_multi_context_coefficient.Rdata"))

write_csv(individual_lm_coefficients_data %>%
            transmute(motif = term, Promoter = promoter, Terminator = terminator, `log2 fold change` = -estimate, p.adj = p.adjust(p.value, method = "fdr")),
          here("supplementary_data_chapter/data/motif_coefficients_and_pvalue.csv"))

```