---
title: "RNAseq vs qPCR comparison"
author: "Sam Haynes"
date: "20/08/2021"
output: pdf_document
---

Now we have QuantSeq and 5PSeq datasets we can compare the abundance of transcripts in these datasets to that determined by qPCR.

Constructs in QuantSeq (Standard RNA-Seq but PolyA Anchored)

Sample Name | Construct Name
------------|---------------
A1 B1       |pRPS3-mCherry-tRPS3_WT
A2 B2 E1 E2 |pRPS3-mCherry-tRPS3_mod0
A3 B3       |pRPS3-mCherry-tRPS3_modB
A4 B4       |pRPS3-mCherry-tRPS3_modE
A5 B5       |pTSA1-mCherry-tTSA1_WT
A6 B6       |pTSA1-mCherry-tTSA1_mod0
A7 B7       |pTSA1-mCherry-tTSA1_modB
A8 B8       |pTSA1-mCherry-tTSA1_modE
A9 B9       |POT1-ccdB
A10 B10     |pRPS3-mCherry-tRPS3_modA
A11 B11     |pTSA1-mCherry-tTSA1_modA
A12 B12     |pPGK1-mCherry-tRPS3_WT
A13 B13     |pPGK1-mCherry-tRPS3_mod0
A14 B14     |pPGK1-mCherry-tRPS3_modA
A15 B15     |pPGK1-mCherry-tRPS3_modB
A16 B16     |pPGK1-mCherry-tRPS3_modE

Constructs in 5PSeq (RNA-Seq of transcripts with 5'Cap removed)

Sample Name | Construct Name
------------|---------------
1dt 2dt     |POT1-ccdB
3dt 4dt     |pRPS3-mCherry-tRPS3_WT
5dt 6dt     |pRPS3-mCherry-tRPS3_mod0
7dt 8dt     |pRPS3-mCherry-tRPS3_modA
9dt 10dt    |pRPS3-mCherry-tRPS3_modB
11dt 12dt   |pRPS3-mCherry-tRPS3_modC
13dt 14dt   |pRPS3-mCherry-tRPS3_modD
15dt 16dt   |pRPS3-mCherry-tRPS3_modE

```{r setup, include=FALSE}
library(tibble)
library(dplyr)
library(tidyr)
library(readr)
library(grid)
library(gridExtra)
library(here)
library(GGally)
library(stringr)
library(DESeq2)
source("/home/dimmest-p/Documents/PhD/3UTR_motifs/motifPrediction/chimera_project_manuscript/raw_data_analysis/code/shared_figure_formatting.R")
```

```{r functions}

read_in_RNAseq <- function(filepath){
  # read in only name and counts
  RNAseq_tibble <- read_tsv(filepath, comment = "#") %>%
    select(1,7)
  
  colnames(RNAseq_tibble) <- c("Geneid", "count")
  
  # replace specific construct Geneid with generic Geneid to allow comparison
  
  construct_URA3_counts <- RNAseq_tibble %>% 
    filter(str_detect(Geneid, "URA3"))
  
  new_name_construct_URA3_counts <- tibble(Geneid = "plasmid_URA3_construct", "count" = construct_URA3_counts %>% pull(2))
  
  construct_RNAseq_tibble <- RNAseq_tibble  %>%
    rows_delete(construct_URA3_counts) %>%
    rows_insert(new_name_construct_URA3_counts)
  
  if(sum(str_detect(RNAseq_tibble$Geneid, "mCh")) > 0){
    construct_counts <- RNAseq_tibble %>% 
    filter(str_detect(Geneid, "mCh"), !str_detect(Geneid, "URA3"))
    
    new_name_construct_counts <- tibble(Geneid = "plasmid_construct", 
                                        "count" = construct_counts %>% pull(2))
    construct_RNAseq_tibble <- construct_RNAseq_tibble %>%
    rows_delete(construct_counts) %>%
    rows_insert(new_name_construct_counts)
  }
  
  # find sample name and add as colname
  colnames(construct_RNAseq_tibble) <- c("Geneid", str_extract(filepath, "(?<=/)[ABEdt0-9]{2,4}(?=/)"))
  
  construct_RNAseq_tibble
}

sample_to_construct_dictionary <- function(filename){
  read_csv(here(filename)) %>%
    mutate(promoter = str_extract(construct_name, "^p[A-Z0-3]+")) %>%
    mutate(terminator = str_extract(construct_name, "t[A-Z0-3]+(?=_)")) %>%
    mutate(mod_label = str_extract(construct_name, "(?<=_)[a-zA-Z0-3]+(?=_)")) %>%
    select(-construct_name) 
}

```

```{r read-RNAseq-counts, message=FALSE, warning=FALSE, results='hide', include=FALSE}
all_QuantSeq_count_files <- list.files(path = here("raw_data_analysis/data/rna_seq/counts/QuantSeq"), pattern = ".txt$", recursive = TRUE, full.names = TRUE)

all_5PSeq_count_files <- list.files(path = here("raw_data_analysis/data/rna_seq/counts/5PSeq"), pattern = ".txt$", recursive = TRUE, full.names = TRUE)

full_QuantSeq_count <- all_QuantSeq_count_files %>%
  lapply(read_in_RNAseq) %>%
  purrr::reduce(full_join, by = "Geneid")

full_5PSeq_count <- all_5PSeq_count_files %>%
  lapply(read_in_RNAseq) %>%
  purrr::reduce(full_join, by = "Geneid")

full_QuantSeq_count_matrix <- full_QuantSeq_count %>% 
  select(-Geneid) %>%
  as.matrix()


full_5PSeq_count_matrix <- full_5PSeq_count %>% 
  select(-Geneid) %>%
  as.matrix()

# replace NAs with 0 
full_QuantSeq_count_matrix[is.na(full_QuantSeq_count_matrix)] <- 0

rownames(full_QuantSeq_count_matrix) <- full_QuantSeq_count %>% 
  pull(Geneid)

full_5PSeq_count_matrix[is.na(full_5PSeq_count_matrix)] <- 0

rownames(full_5PSeq_count_matrix) <- full_5PSeq_count %>% 
  pull(Geneid)

```

```{r sample-name-to-contruct}
sample_to_construct_dictionary_QuantSeq <- sample_to_construct_dictionary("raw_data_analysis/data/rna_seq/sample_to_construct_table_QuantSeq.csv")

sample_to_construct_dictionary_5PSeq <- sample_to_construct_dictionary("raw_data_analysis/data/rna_seq/sample_to_construct_table_5PSeq.csv")

#  swap POT1 empty vector values from NA to "POT"
sample_to_construct_dictionary_QuantSeq <- sample_to_construct_dictionary_QuantSeq %>%
    rows_update(tibble(sample_name = c("A9", "B9"), promoter = "POT1", terminator = "POT1", mod_label = "POT1", rep = c(1,2))) %>%
  mutate(mod_label = factor(mod_label,
                       levels = c(construct_to_label_dictionary_TSA1_RPS3$construct, "POT1"),
                       labels = c(construct_to_label_dictionary_TSA1_RPS3$label, "POT1")))

sample_to_construct_dictionary_5PSeq <- sample_to_construct_dictionary_5PSeq %>%
    rows_update(tibble(sample_name = c("1dt", "2dt"), promoter = "POT1", terminator = "POT1", mod_label = "POT1", rep = c(1,2))) %>%
  mutate(mod_label = factor(mod_label,
                       levels = c(construct_to_label_dictionary_TSA1_RPS3$construct, "POT1"),
                       labels = c(construct_to_label_dictionary_TSA1_RPS3$label, "POT1")))
```

# Could normalising to URA3 cause strange pPGK1 behaviours? Check norm to DESeq2.

```{r create-DESeq2-input}
# tibble holding what condition each column in the count matrix correlates to
coldata_QuantSeq <- tibble(sample_name = colnames(full_QuantSeq_count_matrix)) %>%
  inner_join(sample_to_construct_dictionary_QuantSeq) %>% 
  mutate(col = 1:34) %>%
  mutate(promoter = as.factor(promoter), 
         terminator = as.factor(terminator), 
         mod_label = as.factor(mod_label)) %>%
  column_to_rownames("sample_name") %>%
  filter(promoter != "POT1")

coldata_5PSeq <- tibble(sample_name = colnames(full_5PSeq_count_matrix)) %>%
  inner_join(sample_to_construct_dictionary_5PSeq) %>% 
  mutate(col = 1:16) %>%
  mutate(promoter = as.factor(promoter), 
         terminator = as.factor(terminator), 
         mod_label = as.factor(mod_label)) %>%
  column_to_rownames("sample_name") %>%
  filter(promoter != "POT1")

deseq_count_set_QuantSeq <-  DESeqDataSetFromMatrix(countData = full_QuantSeq_count_matrix[,coldata_QuantSeq %>% pull(col)],
                                           colData = coldata_QuantSeq,
                                           design = ~ promoter * mod_label)

deseq_count_set_5PSeq <-  DESeqDataSetFromMatrix(countData = full_5PSeq_count_matrix[,coldata_5PSeq %>% pull(col)],
                                           colData = coldata_5PSeq,
                                           design = ~ mod_label)
```

```{r run-DESeq}
DESeq_dataset_QuantSeq <- DESeq(deseq_count_set_QuantSeq)

DESeq_dataset_5PSeq <- DESeq(deseq_count_set_5PSeq)
```

```{r view_results}
DESeq_summary_QuantSeq <- results(DESeq_dataset_QuantSeq)

DESeq_summary_5PSeq <- results(DESeq_dataset_5PSeq)

DESeq_summary_QuantSeq[rownames(DESeq_summary_QuantSeq) == "plasmid_construct",]

DESeq_summary_5PSeq[rownames(DESeq_summary_5PSeq) == "plasmid_construct",]

rownames(DESeq_summary_QuantSeq[DESeq_summary_QuantSeq[!is.na(DESeq_summary_QuantSeq$padj),]$padj < 0.05,])

rownames(DESeq_summary_5PSeq[DESeq_summary_5PSeq[!is.na(DESeq_summary_5PSeq$padj),]$padj < 0.05,])

plotMA(DESeq_summary_QuantSeq, ylim=c(-2,2))

plotMA(DESeq_summary_5PSeq, ylim=c(-2,2))
```

```{r compare_whole_genome_expression}
# correlation grouped norm counts
library("pheatmap")

samplesheet_5PSeq<- read_csv(here("raw_data_analysis/data/rna_seq/sample_to_construct_table_5PSeq.csv")) %>%
  mutate(promoter = str_extract(construct_name, "(?<=p)[A-Z0-9]+"))

samplesheet_QuantSeq <- read_csv(here("raw_data_analysis/data/rna_seq/sample_to_construct_table_QuantSeq.csv")) %>%
  mutate(promoter = str_extract(construct_name, "(?<=p)[A-Z0-9]+"))

heatmap_label_QuantSeq_pRPS3 <- as.data.frame(colData(DESeq_dataset_QuantSeq)[,c("promoter","mod_label")]) %>%
  filter(promoter == "pRPS3") %>%
  select(-promoter)

heatmap_label_5PSeq_match_QuantSeq <- as.data.frame(colData(DESeq_dataset_5PSeq)[,c("promoter","mod_label")]) %>%
  filter(mod_label %in% heatmap_label_QuantSeq_pRPS3$mod_label) %>%
  select(-promoter)

log_2_plus_one_QuantSeq <- normTransform(DESeq_dataset_QuantSeq)

log_2_plus_one_5PSeq <- normTransform(DESeq_dataset_5PSeq)

factor_levels_QuantSeq = c("A1", "B1", "E1", "E2", "A2", "B2", "A10", "B10", "A3", "B3", "A4", "B4", "A12", "B12", "A13", "B13", "A14", "B14", "A15", "B15", "A16", "B16", "A5", "B5", "A6", "B6", "A11", "B11", "A7", "B7", "A8", "B8", "A9", "B9")

factor_levels_QuantSeq_single_rep = c("A1", "A2", "A10", "A3", "A4", "A12", "A13", "A14", "A15", "A16", "A5", "A6", "A11", "A7", "A8")

factor_levels_5PSeq_no_poor_sample <- c("3dt", "4dt", "5dt", "6dt", "7dt", "8dt", "9dt", "10dt",                                              "11dt", "13dt", "15dt")

factor_levels_5PSeq_no_poor_sample_single_rep <- c("3dt", "5dt", "7dt", "9dt", "11dt", "13dt", "15dt")

log_2_plus_one_QuantSeq_tibble <- as_tibble(assay(log_2_plus_one_QuantSeq))

sample_correlation_quantseq <- ggplot() + 
    geom_tile(data = as_tibble(cor(log_2_plus_one_QuantSeq_tibble), 
                 rownames = "sample_x") %>% 
         pivot_longer(-sample_x, 
                      names_to = "sample_y", 
                      values_to = "cor") %>% 
         mutate(sample_x = factor(sample_x, factor_levels_QuantSeq), 
                sample_y = factor(sample_y, rev(factor_levels_QuantSeq))),
              aes(x= sample_x, y = sample_y, fill = cor)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8),
        axis.text.y = element_text(size = 8),
        plot.margin = margin()) +
  scale_fill_gradient(high = "grey") +
  coord_equal() +
  labs(x= "", y = "", title = "QuantSeq Sample Correlation") +
  scale_x_discrete(breaks = samplesheet_QuantSeq$sample_name,
                   labels = samplesheet_QuantSeq$type) +
  scale_y_discrete(breaks = samplesheet_QuantSeq$sample_name,
                   labels = samplesheet_QuantSeq$type)
  
  


construct_context_colours_tibble <- tibble(context = factor(c(rep("pRPS3_tRPS3",12),
                                                              rep("pPGK1_tRPS3", 10),
                                                              rep("pTSA1_tTSA1", 10)),
                                                            levels = c("pTSA1_tTSA1",
                                                                       "pPGK1_tRPS3",
                                                                       "pRPS3_tRPS3")))

construct_context_colours_plot <- ggplot() +
  geom_tile(data = construct_context_colours_tibble, aes(x = 1, y = context, fill = context),
            show.legend = FALSE) +
  scale_fill_manual(values = c("pRPS3_tRPS3" = "blue",
                               "pPGK1_tRPS3" = "black",
                               "pTSA1_tTSA1" = "purple")) +
  geom_text(aes(x=1,
                y = c("pTSA1_tTSA1","pPGK1_tRPS3","pRPS3_tRPS3"),
                label = c("pTSA1_tTSA1", "pPGK1_tRPS3", "pRPS3_tRPS3")),
            angle = 90,
            colour = "white") +
  theme_nothing()

log_2_plus_one_5PSeq_tibble <- as_tibble(assay(log_2_plus_one_5PSeq)) %>%
                                     select(factor_levels_5PSeq_no_poor_sample)

sample_correlation_5Pseq <- ggplot(log_2_plus_one_5PSeq_tibble %>%
                                     cor() %>% 
                                     as_tibble(rownames = "sample_x") %>%
         pivot_longer(-sample_x, 
                      names_to = "sample_y", 
                      values_to = "cor") %>% 
         mutate(sample_x = factor(sample_x, factor_levels_5PSeq_no_poor_sample), 
                sample_y = factor(sample_y, rev(factor_levels_5PSeq_no_poor_sample)))) + 
    geom_tile(aes(x= sample_x, y = sample_y, fill = cor)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  coord_equal() +
  labs(x= "", y = "", title = "5PSeq Sample Correlation") +
  scale_x_discrete(breaks = samplesheet_5PSeq$sample_name,
                   labels = samplesheet_5PSeq$type) +
  scale_y_discrete(breaks = samplesheet_5PSeq$sample_name,
                   labels = samplesheet_5PSeq$type)

whole_genome_5PSeq_vs_Quant_Seq<- log_2_plus_one_5PSeq_tibble %>%
  mutate(transcript_name = rownames(log_2_plus_one_5PSeq)) %>%
  pivot_longer(-transcript_name, names_to = "sample_name", values_to = "pseudo_counts") %>%
  inner_join(sample_to_construct_dictionary_5PSeq) %>%
  filter(mod_label %in% sample_to_construct_dictionary_QuantSeq$mod_label) %>%
  mutate(assay = "5PSeq") %>%
  bind_rows(log_2_plus_one_QuantSeq_tibble %>%
              mutate(transcript_name = rownames(log_2_plus_one_QuantSeq)) %>%
              pivot_longer(-transcript_name, names_to = "sample_name", values_to = "pseudo_counts") %>%
              inner_join(sample_to_construct_dictionary_QuantSeq) %>%
              filter(promoter == "pRPS3") %>%
              mutate(assay = "QuantSeq"))

pRPS3_QuantSeq_5PSeq_comparison <- ggplot(whole_genome_5PSeq_vs_Quant_Seq %>%
         select(-sample_name) %>%
         filter(rep<3) %>%
         pivot_wider(names_from = assay, values_from = pseudo_counts)) +
  geom_point(aes(y= `5PSeq`, x = QuantSeq, colour = as.factor(rep))) +
  facet_wrap(~mod_label) +
  coord_equal()

ggplot(whole_genome_5PSeq_vs_Quant_Seq %>%
         filter(transcript_name == "plasmid_construct") %>%
         select(-sample_name) %>%
         group_by(mod_label, assay) %>%
         summarise(pseudo_counts = mean(pseudo_counts)) %>%
         pivot_wider(names_from = assay, values_from = pseudo_counts)) +
  geom_point(aes(y= `5PSeq`, x = QuantSeq, colour = mod_label)) +
  coord_equal() +
  scale_colour_manual(values = RPS3_TSA1_colour_scheme)


ggpairs_QuantSeq_vs_5PSeq <- whole_genome_5PSeq_vs_Quant_Seq %>% 
  filter(rep == 1) %>% 
  arrange(mod_label) %>% 
  mutate(mod_label = str_remove(mod_label, "mod_"),
         assay = str_replace(assay, "5PSeq", "5PS"),
         assay = str_replace(assay, "QuantSeq", "QS")) %>%
  unite(source, assay, mod_label) %>% 
  select(-sample_name, -rep, -promoter, -terminator, -type) %>% 
  pivot_wider(names_from = source, values_from = pseudo_counts) %>% 
  select(-transcript_name) %>% 
  ggpairs(upper = list(continuous = wrap("cor", size = 3)))

```


```{r direct_compare_Quant_and_5PSeq_construct_abund}
median_normalising_gene_QuantSeq <- full_QuantSeq_count %>%
  filter(str_detect(Geneid, c("plasmid_U|YML028W|YNL178W|YCR012W"))) %>%
  pivot_longer(-Geneid, names_to = "sample_name", values_to = "counts") %>%
  group_by(sample_name) %>%
  summarise(norm_value = median(counts))
  
qPCR_like_normalised_QuantSeq <- full_QuantSeq_count %>%
  filter(str_detect(Geneid, "plasmid_c"))  %>%
  pivot_longer(-Geneid, names_to = "sample_name", values_to = "counts") %>%
  inner_join(median_normalising_gene_QuantSeq) %>%
  mutate(norm_count = counts / norm_value) %>%
  inner_join(sample_to_construct_dictionary_QuantSeq) %>%
  group_by(promoter, mod_label) %>%
  summarise(norm_count = mean(norm_count)) %>%
  ungroup() %>%
  filter(!is.na(norm_count))  %>%
  mutate(`RNA-Seq` = "QuantSeq")

median_normalising_gene_5PSeq <- full_5PSeq_count %>%
  filter(str_detect(Geneid, c("plasmid_U|YML028W|YNL178W|YCR012W"))) %>%
  pivot_longer(-Geneid, names_to = "sample_name", values_to = "counts") %>%
  group_by(sample_name) %>%
  summarise(norm_value = median(counts)) %>%
  filter(norm_value > 0)
  
qPCR_like_normalised_5PSeq <- full_5PSeq_count %>%
  filter(str_detect(Geneid, "plasmid_c"))  %>%
  pivot_longer(-Geneid, names_to = "sample_name", values_to = "counts") %>%
  inner_join(median_normalising_gene_5PSeq) %>%
  mutate(norm_count = counts / norm_value) %>%
  inner_join(sample_to_construct_dictionary_5PSeq) %>%
  group_by(promoter, mod_label) %>%
  summarise(norm_count = mean(norm_count)) %>%
  ungroup() %>%
  filter(!is.na(norm_count)) %>%
  mutate(`RNA-Seq` = "5PSeq")

RNASeq_construct_comparison <- qPCR_like_normalised_QuantSeq %>%
  bind_rows(qPCR_like_normalised_5PSeq) %>%
  filter(promoter == "pRPS3") %>%
  select(-promoter) %>%
  pivot_wider(names_from = `RNA-Seq`, values_from = norm_count) %>%
  filter(!is.na(QuantSeq))

RNASeq_comparison_R_values <- RNASeq_construct_comparison %>%
  summarise(r_cor = cor(QuantSeq, `5PSeq`)) %>%
  mutate(r_cor_label = paste(TeX(paste0("R = ", signif(r_cor,digits = 3)))))

QuantSeq_5PSeq_construct_abund <- ggplot(RNASeq_construct_comparison) +
  geom_point(aes(x = QuantSeq, y = `5PSeq`, colour = mod_label)) +
  scale_colour_manual(values = RPS3_TSA1_colour_scheme[c("WT", "mod_HTH", "mod_NAA", "mod_NNN", "mod_NTN")]) +
  geom_abline(slope = 1, intercept = 0) +
  scale_y_continuous(breaks = c(1,2)) +
  theme(legend.position = "bottom",
        legend.margin=margin(t = 0, unit='cm'),
        legend.justification = c(0.5,0)) +
  guides(colour = guide_legend(ncol = 2,
                               title = "Construct",
                               title.position = "top",
                               title.hjust = 0.5)) + 
  geom_text(aes(label = r_cor_label), 
            data = RNASeq_comparison_R_values, 
            x = 4.2, y = 1, 
            size = 3, hjust = 1, vjust = 0)
```


```{r output-cor-graphs, eval = FALSE}
ggsave2(here("supplementary_data_chapter/figures/sequencing_sample_correlation.png"),
        plot_grid(sample_correlation_quantseq,
                                         plot_grid(sample_correlation_5Pseq,
                                                   QuantSeq_5PSeq_construct_abund,
                                                   labels = c("B", "C"),
                                                   scale = c(1.35,0.8),
                                                   axis = "bottom",
                                                   hjust = c(1, -3),
                                                   align = "hv"), 
                                         ncol = 1,
                                         scale = c(1,0.8),
                                         labels = c("A", ""),
                                         hjust = c(-3.7, 0)),
        width = fig_width_2column,
        height = 210,
        units = "mm",
        dpi = fig_dpi,
        bg = "white")

ggsave2(here("supplementary_data_chapter/figures/pRPS3_QuantSeq_5PSeq_comparison.png"),
        ggpairs_QuantSeq_vs_5PSeq + 
          theme(axis.text = element_text(size = 8),
                strip.text.x = element_text(size = 8),
                strip.text.y = element_text(size = 8)),
        width = fig_width_2column,
        height = fig_width_2column,
        units = "mm",
        dpi = fig_dpi,
        bg = "white")
```